---
title: "Combined proteome + phosphoproteome analysis"
author: "Sun et al."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)
library(reshape2)
library(ggrepel)
library(gridExtra)
library(ggpubr)
library(cowplot)
```

```{r load-data}
# DE results: protein log2FC (WT/rim15KO) per condition × timepoint
logFC_ttest_pG <- read_csv("./tables/DE_proteins_byGenotype.csv",
                           show_col_types = FALSE)

# DE results: phosphosite log2FC (WT/rim15KO) per condition × timepoint
logFC_ttest_pS <- read_csv("./tables/DE_pSites_byGenotype.csv",
                           show_col_types = FALSE)

# Normalised protein ratios (wide: one row per protein × sample)
master <- read_csv("./tables/master_tidy_protein.csv",
                   show_col_types = FALSE) %>%
  mutate(samplingTime = as.numeric(substr(samplingTime, 2, nchar(samplingTime))))

# ANCOVA results — used to define sig protein lists for trajectory plots
anova_genotype         <- read_csv("./tables/anova_genotype.csv",
                                   show_col_types = FALSE)
anova_nutrient         <- read_csv("./tables/anova_nutrient.csv",
                                   show_col_types = FALSE)
anova_genotype_nutrient <- read_csv("./tables/anova_genotype_nutrient.csv",
                                    show_col_types = FALSE)
```

```{r define-sig-lists}
# Proteins with significant genotype × time interaction (q < 0.05, normed_ratio)
sig_genotype <- anova_genotype %>%
  filter(term == "samplingTime:genotypeWT", data == "normed_ratio", qvalue < 0.05)

# Proteins with significant nutrient × time interaction (q < 0.05, normed_ratio)
sig_nutrient <- anova_nutrient %>%
  filter(grepl("samplingTime", term) & grepl("nutrient", term),
         data == "normed_ratio", qvalue < 0.05)

# Proteins with significant 3-way interaction (q < 0.05, normed_ratio)
sig_3way <- anova_genotype_nutrient %>%
  filter(term == "samplingTime:genotype:nutrient",
         data == "normed_ratio", qvalue < 0.05)
```

## Phosphosite vs protein log2FC scatter

Scatter plot of protein-level log2FC (x) against phosphosite log2FC (y). Red points
highlight phosphosites with significant change (q < 0.05) that exceed the parent
protein's fold change — i.e. sites whose change is not simply a reflection of altered
protein abundance.

```{r scatter-phospho-vs-protein, fig.width=10, fig.height=8}
logFC_ttest <- left_join(
  logFC_ttest_pS,
  logFC_ttest_pG,
  by     = c("Protein names", "Gene names", "Fasta headers", "nutrient", "time"),
  suffix = c(".phospho", ".protein")
)

p_scatter <- ggplot(data = logFC_ttest,
                    aes(x = log2FC.protein, y = log2FC.phospho)) +
  geom_point(size = 0.8, alpha = 0.5, color = "#a3a3a3") +
  geom_point(data = filter(logFC_ttest,
                           log2FC.phospho > 0,
                           qvalue.phospho < 0.05,
                           log2FC.phospho > log2FC.protein),
             color = "#FF3333", size = 1, alpha = 0.7) +
  geom_point(data = filter(logFC_ttest,
                           log2FC.phospho < 0,
                           qvalue.phospho < 0.05,
                           log2FC.phospho < log2FC.protein),
             color = "#3399FF", size = 1, alpha = 0.7) +
  geom_text_repel(
    data  = filter(logFC_ttest,
                   log2FC.phospho > 0, qvalue.phospho < 0.05,
                   log2FC.phospho > log2FC.protein),
    aes(label = pSite), size = 2, max.overlaps = 15) +
  geom_text_repel(
    data  = filter(logFC_ttest,
                   log2FC.phospho < 0, qvalue.phospho < 0.05,
                   log2FC.phospho < log2FC.protein),
    aes(label = pSite), size = 2, max.overlaps = 15) +
  geom_vline(xintercept = 0, color = "#cccccc", linetype = "dashed") +
  geom_hline(yintercept = 0, color = "#cccccc", linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, color = "#888888", linetype = "dotted") +
  xlab(expression(log[2] * " (WT / rim15Δ) - protein")) +
  ylab(expression(log[2] * " (WT / rim15Δ) - phosphorylation site")) +
  facet_grid(nutrient ~ time, scale = "free") +
  theme_classic(base_size = 11) +
  theme(legend.position = "none",
        strip.background = element_blank())

p_scatter

ggsave("./plots/phospho_vs_protein_logFC.pdf",
       plot = p_scatter, width = 10, height = 7, useDingbats = FALSE)
```

## Trajectory plots for significantly DE proteins

The next three chunks generate multi-page PDFs of protein trajectories for proteins
significant by genotype, nutrient, and the 3-way interaction respectively.
Set `eval=TRUE` to regenerate; they are slow for large protein lists.

```{r traj-sig-genotype, eval=FALSE}
master_long <- master %>%
  pivot_longer(c(normalized_ratio, normalized_intensity),
               names_to = "variable", values_to = "value") %>%
  filter(!is.na(value), is.finite(value))

protein_list <- unique(sig_genotype$`Protein IDs`)
plot_list_genotype <- list()

for (cond in c("C", "P")) {
  for (i in seq_along(protein_list)) {
    protein <- master_long %>%
      filter(nutrient == cond, `Protein IDs` == protein_list[i])
    if (nrow(protein) == 0) next

    p <- ggline(protein, x = "samplingTime", y = "value",
                color = "genotype", facet.by = "variable",
                add = c("mean_se", "dotplot"),
                palette = c("#00AFBB", "#E7B800")) +
      facet_wrap(~variable, scale = "free", ncol = 2) +
      ggtitle(paste0(cond, " — ", unique(protein$`Gene names`))) +
      scale_y_log10()

    plot_list_genotype[[length(plot_list_genotype) + 1]] <- p
  }
}

ggsave("./plots/sigDE_by_genotype.pdf",
       plot = marrangeGrob(plot_list_genotype, nrow = 6, ncol = 1),
       width = 12, height = 14)
```

```{r traj-sig-nutrient, eval=FALSE}
master_long <- master %>%
  pivot_longer(c(normalized_ratio, normalized_intensity),
               names_to = "variable", values_to = "value") %>%
  filter(!is.na(value), is.finite(value))

protein_list <- unique(sig_nutrient$`Protein IDs`)
plot_list_nutrient <- list()

for (gt in c("WT", "rim15KO")) {
  for (i in seq_along(protein_list)) {
    protein <- master_long %>%
      filter(genotype == gt, `Protein IDs` == protein_list[i])
    if (nrow(protein) == 0) next

    p <- ggline(protein, x = "samplingTime", y = "value",
                color = "nutrient", facet.by = "variable",
                add = c("mean_se", "dotplot"),
                palette = c("#ff9500", "#8300ab")) +
      facet_wrap(~variable, scale = "free", ncol = 2) +
      ggtitle(paste0(gt, " — ", unique(protein$`Gene names`))) +
      scale_y_log10()

    plot_list_nutrient[[length(plot_list_nutrient) + 1]] <- p
  }
}

ggsave("./plots/sigDE_by_nutrient.pdf",
       plot = marrangeGrob(plot_list_nutrient, nrow = 6, ncol = 1),
       width = 12, height = 14)
```

```{r traj-sig-3way, eval=FALSE}
master_long <- master %>%
  pivot_longer(c(normalized_ratio, normalized_intensity),
               names_to = "variable", values_to = "value") %>%
  filter(!is.na(value), is.finite(value))

protein_list <- unique(sig_3way$`Protein IDs`)
plot_list_3way <- list()

for (gt in c("WT", "rim15KO")) {
  for (i in seq_along(protein_list)) {
    protein <- master_long %>%
      filter(genotype == gt, `Protein IDs` == protein_list[i])
    if (nrow(protein) == 0) next

    p <- ggline(protein, x = "samplingTime", y = "value",
                color = "nutrient", facet.by = "variable",
                add = c("mean_se", "dotplot"),
                palette = c("#ff9500", "#8300ab")) +
      facet_wrap(~variable, scale = "free", ncol = 2) +
      ggtitle(paste0(gt, " — ", unique(protein$`Gene names`))) +
      scale_y_log10()

    plot_list_3way[[length(plot_list_3way) + 1]] <- p
  }
}

ggsave("./plots/sigDE_3way.pdf",
       plot = marrangeGrob(plot_list_3way, nrow = 6, ncol = 1),
       width = 12, height = 14)
```
