---
title: "Figure 1"
author: "Siyu Sun"
date: "9/9/2020"
output: html_document
---

#Load essential packages
```{r,cache=TRUE}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(gplots) 
library(ggrepel)
library(ggformula)
library(mosaicData)
library(tidyverse)
library(tidyr)
library(readr)
library(cowplot)
library(scales)
library(broom)
library(dplyr)
```

#read in protein groups that have been converted into .csv
#no biological replicate on 190621, cells were cultured with C (0.06%), N, P supplemented with natural version of lys and arg with the expected concentration. 

#Fig1A
##this data only contains cells cultured in heavy and medium aa supplimented medium, with 0.1mM of Arg and Lys
```{r}
growth_long <- read_csv("./growth_long.csv")

df <- growth_long
model_rr <- df %>%
  #remove the redundance readings for cell counts
  #select(-rep_sub) %>%
  #unite(rep_rep, rep, rep_sub) %>%
  group_by(strain, nutrient) %>%
  nest() %>%
  #fit gc
  mutate(pr.model = map(data, ~SummarizeGrowth(.$time, .$counts))) %>%
  mutate(predict = map(pr.model, ~predict(.$model))) %>%
  select(-pr.model) %>%
  #unnest the tibble and combine the predicted data with raw reading
  unnest(cols = c(data, predict))

#plot growth curve for the whole growth stage
model_rr %>%
  ggplot(aes(x = time, y = counts, col=strain)) +
  geom_point(aes(shape = as.character(replicate)), size = 1.5) +
  geom_line(aes(y=predict, col = strain), size = 1.5) +
  facet_wrap(~nutrient, scale = "free") +
  scale_color_manual(values=c("#00afbb", "#e7b800")) +
  scale_fill_manual(values=c("#00afbb", "#e7b800")) +
  scale_y_log10() +
  theme_classic(base_size = 18) +
  theme(legend.position="bottom")+
  labs(x="Hours post nutrient removal", y = "log10 (cell counts/ml)")
ggsave("gc_rr_long_model.pdf",width = 10, height = 4) 

#growth rate difference between wt and rim15 under different environmental condition
summarized <- df %>%
  #select(-rep_sub) %>%
  group_by(strain, time, nutrient) %>%
  summarize(ave = mean(counts), sd = sd(counts)) 

#plot for the short timeframe
summarized %>%
  ggplot(aes(x = time, y = ave, col=strain)) +
  geom_line(size = 1.5) +
  geom_point(aes(shape = strain), size = 1) +
  geom_errorbar(aes(ymin = ave - sd, ymax = ave + sd, col = strain), width = 1, size = 1)+
  scale_color_manual(values=c("#00afbb", "#e7b800")) +
  scale_fill_manual(values=c("#00afbb", "#e7b800")) +
  facet_wrap(~nutrient, scale = "free") +
  #scale_y_continuous(trans = "log10", limits = c(1,100)) +
  scale_y_log10() +
  theme_classic(base_size = 18)+
  theme(legend.position="bottom")+
  labs(x="Hours post nutrient removal", y = "log10 (cell counts/ml)")
ggsave("Fig1A.pdf",width = 12, height = 4.5) 
```

#Fig1B
## viability
```{r}
df<- read_csv("./viability.csv")
#convert the viability percentage into numeric for plotting
df$viability = round(as.numeric(gsub("[\\%,]", "", df$viability)),digits = 1)

#plot viability difference between wt and rim15 under each condtion
#caculate the viability in relative to T0:
df_rela <- df %>% 
  select(strain, time, nutrient, replicate, viability) %>%
  dplyr::filter(strain %in% c("wt","rim15")) %>%
  spread(key = time, value = viability) %>%
  dplyr::mutate(`0` = 100) %>%
  #dplyr::filter(Media !="SILAC_L") %>%
  dplyr::mutate(T0 = `0`/`0`,T6 = `6`/`0`,T16 = `16`/`0`,T30 = `30`/`0`,T168 = `168`/`0`,T264 = `264`/`0`) %>%
  melt(key = "replicate") %>%
  dplyr::filter(strain %in% c("wt","rim15")) %>%
  #dplyr::filter(variable %in% c("T0","T6","T16","T30","T168", "T264")) %>%
  mutate(variable = gsub("[^0-9.-]", "", variable)) %>%
  #dplyr::filter(SamplingTime %in% c("0","9","20","46")) %>%
  dplyr::select(strain, nutrient, variable, value) %>%
  #dplyr::filter(Media != "SILAC_L") %>%
  dplyr::rename(time = variable) %>%
  #dplyr::mutate(nutrient = case_when(
  #   nutrient == "C(0.01%)+L" ~ "C", 
  #   nutrient == "N-dep+L" ~ "N",
  #   nutrient == "P-dep+L" ~ "P")) %>%
  dplyr::mutate(time = as.numeric(time)) 

combine <- left_join(df, df_rela , by = c("strain", "time","nutrient")) %>%
  drop_na()

combine$strain <- factor(combine$strain, levels = c("wt","rim15"))

combine <- combine %>%
  group_by(strain, time, nutrient) %>%
  summarise(gc.mean = mean(counts),gc.sd = sd(counts), via.mean = mean(value), via.sd = sd(value))

scaleFactor <- max(combine$gc.mean) / max(combine$via.mean)

ggplot(df, aes(x=time, y = viability, color = strain, group = strain)) +
  geom_point(aes(y=viability, group=strain, color=strain, 
                 shape = "circle"), size = 2) +
  geom_smooth(aes(y=viability , alpha = 0.3, group=strain, color=strain, linetype = "solid"), stat="lm", se=F, formula = y ~ x, size = 1.5) +
  stat_cor(aes(group=strain, color=strain), label.y = 25, size = 5) +
  stat_regline_equation(aes(group=strain, color=strain), label.y = 20, size = 5) +#this means at 30th unit regresion line equation will be shown
  #stat_poly_eq(formula = y ~ x, 
   #            aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
    #           parse = TRUE, size = 5) +   
  scale_y_continuous(name="Viability")+
  theme_classic(base_size = 18) +
  scale_color_manual(values=c("#00afbb", "#e7b800")) +
  scale_fill_manual(values=c("#00afbb", "#e7b800")) +
  facet_wrap(~nutrient, scale = "free") +
  theme(legend.position="bottom")
ggsave("Fig1B.pdf",width = 12, height = 4.5) 
```

#Fig1C
##Bud index and gc for short time frame
```{r}
budex <- read_csv("./budindex_short.csv")
growth <- read_csv("./growth_short.csv")
#caculate the mean and std
mu <- budex %>%
  group_by(strain, time, nutrient) %>%
  summarize(mean = mean (`bud index`), ste = sd(`bud index`)/sqrt(length(`bud index`)))

#Combine growth and index data into the same table and show them on the same plot
combine <- left_join(mu, growth %>% rename(time = SamplingTime), by = c("strain", "time","nutrient"))
combine$strain <- factor(combine$strain, levels = c("wt","rim15"))

scaleFactor <- max(combine$cellcounts) / max(combine$mean)

ggplot(combine, aes(x=time)) +
  geom_point(aes(y=cellcounts, group=strain, color=strain, shape = "circle") , size = 2) +
  geom_point(aes(y=mean * scaleFactor, group=strain, color=strain,alpha = 0.3, shape = "trangle"),size = 2) + 
  geom_line(aes(y=cellcounts, group=strain, color=strain, linetype = "dashed"), size = 1.5) +
  geom_line(aes(y=mean * scaleFactor, group=strain, color=strain, linetype = "solid"), size = 1.5) +
  geom_errorbar(aes(ymin=mean * scaleFactor - ste*scaleFactor, ymax=mean * scaleFactor + ste*scaleFactor, color=strain), width=.1, size = 1, position=position_dodge(0.1)) +
  scale_y_continuous(name="Cell concnetraion (solid line)", sec.axis=sec_axis(~./scaleFactor, name="Bud index (dashed line)")) + #trans = "log10", 
  geom_vline(xintercept = c(0,6,16,30), linetype = "dotted", size = 0.7) +
  scale_color_manual(values=c("#e7b800", "#00afbb")) +
  scale_fill_manual(values=c("#e7b800", "#00afbb")) +
  facet_grid(~nutrient, scale = "free") +
  #theme_classic(base_size = 18, ) +
  theme_grey(base_size = 18) +
  theme(
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red"),
    legend.position="bottom"
  )
ggsave("Fig1C.pdf",width = 12, height = 4.5)  
```


#Fig1D
##incorperation efficiency between prototrophic wt and lysarg auxothrope
```{r}

wt_M <- read_csv("./proteinGroups_pro_auxo_M.csv") %>%
  #only keep the protein id columns and columns with column names contain "proteome" and "Intensity"
  dplyr::select(`Protein IDs`, contains("Intensity")) %>%
  dplyr::select(`Protein IDs`, contains(c("4","5"))) %>%
  #remove rows with all 0 readings in every column
  filter_at(vars(-`Protein IDs`), any_vars(. != 0)) 

wt_H <- read_csv("./proteinGroups_pro_auxo_H.csv") %>%
  #only keep the protein id columns and columns with column names contain "proteome" and "Intensity"
  dplyr::select(`Protein IDs`, contains("Intensity")) %>%
  dplyr::select(`Protein IDs`, contains(c("4","5"))) %>%
  #remove rows with all 0 readings in every column
  filter_at(vars(-`Protein IDs`), any_vars(. != 0))


```

##incoperation efficiency if rim15
```{r}
#read in rim15KO data
protein_rim15 <- read_csv("./proteinGroups_rim15_HM.csv")

df <- protein_rim15 %>%
  #only keep the protein id columns and columns with column names contain "proteome" and "Intensity"
  dplyr::select(`Protein IDs`, contains("Intensity")) %>%
  dplyr::select(`Protein IDs`, contains(c("1","2"))) %>%
  #remove rows with all 0 readings in every column
  filter_at(vars(-`Protein IDs`), any_vars(. != 0)) 

#reshape the table
df_rs <- df %>%
  melt() %>%
  separate(variable, c("signal","channel","sampleNum"), sep = " ") %>%
  drop_na()
  
#replace NA values in sample Num with corresponding channel %>%
df_rs$sampleNum[is.na(df_rs$sampleNum)] <- df_rs$channel[is.na(df_rs$sampleNum)]

df_rs <- df_rs %>% 
  #replace channel by total if channel is 1 or 2
  dplyr::mutate(channel = ifelse(channel %in% c(1,2), "total", channel)) %>%
  #replace sample number 1 by rim15-M and sample number 2 by rim15-H
  dplyr::mutate(sampleNum = ifelse(sampleNum == 1, "rim15-M", "rim15-H")) %>%
  #add 1 to the intensity of each protein for later log transformation 
  dplyr::mutate(value = value + 1) %>%
  dplyr::select(-signal) 

#look at the ratio of total intensity for each channel
summed_channel_wise <- df_rs %>%
  filter(channel != "total") %>%
  dplyr::group_by(channel, sampleNum) %>%
  dplyr::summarize(summed = sum(value), mean = mean(value))

summed_sample_wise <- summed_channel_wise %>%
  dplyr::group_by(sampleNum) %>%
  dplyr::summarize(total = sum(summed))

summary_tb <- left_join(summed_channel_wise,summed_sample_wise, by = "sampleNum") %>%
  mutate(fraction = summed/total)

#make a barplot for the total incoperation efficiency in each sample
total_bar <- ggplot(summary_tb, aes(fill=channel, y=fraction, x=sampleNum)) + 
  geom_bar(stat="identity") +
  scale_color_manual(values=c("#4b006e", "#e5a3ff", "#ae00ff"))+
  scale_fill_manual(values=c("#4b006e", "#e5a3ff", "#ae00ff")) +
  #facet_wrap(~sampleNum, scale = "free") +
  coord_polar("y", start=0) +
  theme(legend.position="bottom")+
  theme_bw(base_size = 9) +
  xlab("") +
  ylab("")

#select heavy labeled samples
df_H <- df %>%
  dplyr::select(`Protein IDs`, contains(c("2"))) %>%
  mutate(`rim15-H` = `Intensity H 2`/`Intensity 2`) %>%
  drop_na() %>%
  select(`Protein IDs`,`rim15-H`) 

df_M <- df %>%
  dplyr::select(`Protein IDs`, contains(c("1"))) %>%
  mutate(`rim15-M` = `Intensity M 1`/`Intensity 1`) %>%
  drop_na() %>%
  select(`Protein IDs`,`rim15-M`)

#combine medium and heavy samples
frac<-left_join(df_M,df_H, by = "Protein IDs")

p <- frac %>% 
  dplyr::select(`rim15-M`, `rim15-H`) %>%
  melt() %>%
  ggplot(aes(x=value, col= variable, fill=variable))+
  theme_bw(base_size=6)+
  theme(axis.text.x=element_text(angle=90, size=6), panel.grid.major = element_blank())+
  #geom_vline(x=0.98, linetype="dashed")+
  geom_density(alpha = 0.7)+
  xlim(0.95,1)+
  scale_color_manual(values=c("#ae00ff", "#4b006e"))+
  scale_fill_manual(values=c("#ae00ff", "#4b006e"))+
  labs(x="% of labeled proteins", y = "Density")+
  theme_classic(base_size = 22)

ggdraw(p + theme_half_open(11)) +
  draw_plot(total_bar, .05, .15, 0.7, 0.65) 

ggsave("./incoperation_rim15.pdf", width=7,height=5)
```