---
title: "Cross-dataset comparison of Rim15-dependent phosphosites (R1.3 / R2.5)"
output: html_document
---

Both reviewers asked for a quantitative overlap of our Rim15-dependent
phosphosites with published Rim15 / PP2A-Cdc55 / quiescence phosphoproteomes:
an UpSet plot (plus optional 3-set Venn) and a supplementary membership table.

Datasets (supplement files in `data/published/`, kept locally / gitignored;
sources listed in `data/published/README.md`):

- **This study** — significant sites from `tables/DE_pSites_byGenotype.csv`
  (site key already `GENE_position`, e.g. `STE2_329`).
- **Li et al. 2019** (`data/published/li2019/`) — Rim15-dependent sites.
- **Dokladal et al. 2021** (`data/published/dokladal2021/`) — TORC1/Rim15 axis.
- **Baro et al. 2018** (`data/published/baro2018/`) — PP2A-Cdc55.
- **Plank et al. 2020** (`data/published/plank2020/`) — quiescence phospho.

Harmonization plan (the error-prone step is residue numbering — different
studies used different reference proteomes; verify a handful of well-known
sites, e.g. Igo1 S64, by hand before trusting the overlap):

1. Map every protein to its SGD systematic ORF name (use org.Sc.sgd.db or the
   SGD alias table); site key = `ORF_residueLetter_position`.
2. Compute overlap at two resolutions: (i) protein-level, (ii) exact-site-level.
3. UpSet plot (UpSetR) + full membership table
   -> `revision3/output/phosphosite_overlap_upset.pdf`,
   `revision3/output/phosphosite_membership.csv`.
4. Define the consensus set (sites in >= 2 published datasets + ours) and the
   novel set (ours only); these numbers feed the Discussion rewrite.

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
# library(UpSetR)
# library(org.Sc.sgd.db)

dir.create("./revision3/output", showWarnings = FALSE, recursive = TRUE)

# --- This study: significant Rim15-dependent sites --------------------------
ours <- read_csv("./tables/DE_pSites_byGenotype.csv") %>%
  filter(qvalue <= 0.05) %>%
  distinct(`Gene names`, pSite)

# --- TODO: identify the site-level table within each supplement -------------
# Inspect sheets first, e.g.:
# excel_sheets("./data/published/li2019/mmc2.xlsx")
# Each dataset needs: gene/ORF, residue letter, position, and the study's
# significance call for Rim15- (or Cdc55-) dependence.
```

```{r session}
sessionInfo()
```
