---
title: "Cross-dataset comparison of Rim15-dependent phosphosites (R1.3 / R2.5)"
output: html_document
---

Both reviewers asked for a quantitative overlap of our Rim15-dependent
phosphosites with published Rim15 / PP2A-Cdc55 / quiescence phosphoproteomes:
an UpSet plot (plus optional 3-set Venn) and a supplementary membership table.

Datasets (supplement files in `data/Publications for comparative analysis/`, kept
locally / gitignored; sources listed in `data/Publications for comparative analysis/README.md`):

- **This study** — significant sites from `tables/DE_pSites_byGenotype.csv`
  (q <= 0.05 in any nutrient x time contrast); residue letter and leading
  UniProt accession recovered from `data/Phospho (STY)Sites.csv`.
- **Li et al. 2019** (`Phosphoproteome/li2019/mmc5.xlsx`, "Phospho Regulation Network") —
  phosphosites regulated in the rim15 deletion strain; multi-site keys
  (e.g. `POL12_pT153pS168`) are exploded into individual residues.
- **Dokladal et al. 2021** (`Phosphoproteome/dokladal2021/mmc3.xlsx`, "Exclusive") — high
  confidence Rim15 target sites (rim15∆ block, columns 24-29); residue letter
  taken from the center of the 15-mer sequence motif.
- **Baro et al. 2018** (`Phosphoproteome/baro2018/Additional file 3.xlsx`) — hyperphosphorylated
  peptides of putative PP2A-Cdc55 regulated proteins; exact positions parsed
  from the Modifications column (e.g. `S9(Phospho)`), mapped to absolute
  coordinates by locating the peptide in the UniProt protein sequence.
- **Plank et al. (Cdc55-AID)** (`Phosphoproteome/plank2020/pmic13349-sup-0002-tables1.xlsx`) —
  acute Cdc55 depletion (auxin vs mock). Two caveats. (a) At the study's own
  FDR threshold essentially nothing is significant (min pAdj = 0.0496, their
  `below0.05FDR` flag is FALSE for every peptide) — 10 min of Cdc55 depletion
  barely moves the phosphoproteome — so we use *nominal* p <= 0.05 with
  increased phosphorylation (log2 aux/mock > 0, the substrate-consistent
  direction); interpret this set accordingly. (b) The supplement is a MaxQuant
  modification-specific-peptide table without site positions, so exact sites
  are assigned only where the number of S/T/Y in the peptide equals the
  phospho count; otherwise the protein contributes to the protein-level
  overlap only.

Harmonization (the error-prone step is residue numbering — different studies
used different reference proteomes; Igo1 S64 is checked by hand below):

1. Every protein is mapped to its SGD systematic ORF via org.Sc.sgd.db
   (UniProt accession where available, otherwise gene symbol with
   GENENAME/COMMON/ALIAS fallbacks). Site key = `ORF_residueLetter+position`.
2. Overlap at two resolutions: (i) protein-level, (ii) exact-site-level.
3. UpSet plots (UpSetR) + full membership tables
   -> `revision3/output/phosphosite_overlap_upset.pdf` (site-level),
   `revision3/output/phosphoprotein_overlap_upset.pdf` (protein-level),
   `revision3/output/phosphosite_membership.csv`,
   `revision3/output/phosphoprotein_membership.csv`.
4. Consensus set = our sites recovered in >= 2 published datasets; novel set
   = our sites in none of them. Headline numbers printed at the end feed the
   Discussion rewrite and the response letter.

Peptide-to-protein mapping uses the UniProt reference proteome UP000002311
(`data/Publications for comparative analysis/UP000002311_559292.fasta`, downloaded 2026-08-24).

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
library(data.table)      # fread with column selection for the wide MaxQuant table
library(Biostrings)
library(org.Sc.sgd.db)
library(UpSetR)

repo_root <- "/Users/david/Projects/Proteome_Rim15_Quiescence"
out_dir   <- file.path(repo_root, "revision3/output")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# ORF-shaped identifiers (Y-ORFs incl. dubious suffixes, Q-ORFs) pass through
is_orf <- function(x) grepl("^(Y[A-P][LR]\\d{3}[WC](-[A-Z])?|Q\\d{4})$", x)

# Map a vector of gene symbols to systematic ORF names, trying successively
# looser keytypes. Returns NA where nothing maps.
symbol_to_orf <- function(symbols) {
  key <- toupper(trimws(symbols))
  orf <- ifelse(is_orf(key), key, NA_character_)
  for (kt in c("GENENAME", "COMMON", "ALIAS")) {
    if (!kt %in% keytypes(org.Sc.sgd.db)) next
    todo <- unique(key[is.na(orf)])
    todo <- todo[!is.na(todo) & todo != ""]
    if (length(todo) == 0) break
    m <- tryCatch(
      AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = kt,
                            columns = "ORF"),
      error = function(e) NULL)
    if (is.null(m)) next
    m <- m %>% filter(!is.na(ORF)) %>% distinct(across(1), ORF) %>%
      group_by(across(1)) %>% dplyr::slice(1) %>% ungroup()
    lookup <- setNames(m$ORF, m[[1]])
    orf[is.na(orf)] <- lookup[key[is.na(orf)]]
  }
  orf
}

uniprot_to_orf <- function(accessions) {
  acc <- sub("-\\d+$", "", trimws(accessions))   # strip isoform suffix
  todo <- unique(acc[!is.na(acc) & acc != ""])
  m <- AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = "UNIPROT",
                             columns = "ORF") %>%
    filter(!is.na(ORF)) %>% distinct(UNIPROT, ORF) %>%
    group_by(UNIPROT) %>% dplyr::slice(1) %>% ungroup()
  setNames(m$ORF, m$UNIPROT)[acc]
}

site_key <- function(orf, aa, pos) paste0(orf, "_", aa, pos)
```

```{r fasta}
pub_dir <- file.path(repo_root, "data/Publications for comparative analysis")
proteome <- readAAStringSet(file.path(pub_dir, "UP000002311_559292.fasta"))
# index by accession (header: sp|P17065|SEC2_YEAST ...)
names(proteome) <- sub("^..\\|([^|]+)\\|.*$", "\\1", names(proteome))
proteome_chr <- setNames(as.character(proteome), names(proteome))

# Locate a peptide in its protein; returns start position (NA if no match).
# Peptides matching more than once are rare; we take the first match and flag.
locate_peptide <- function(peptide, accession) {
  seqs <- proteome_chr[accession]
  starts <- mapply(function(p, s) {
    if (is.na(s)) return(NA_integer_)
    r <- gregexpr(p, s, fixed = TRUE)[[1]]
    if (r[1] == -1) NA_integer_ else as.integer(r[1])
  }, toupper(peptide), seqs, USE.NAMES = FALSE)
  starts
}
```

# This study

```{r ours}
de <- read_csv(file.path(repo_root, "tables/DE_pSites_byGenotype.csv"), show_col_types = FALSE)
# Filter explicitly to C and P starvation (N starvation was not profiled
# in the phosphoproteomics; this makes the scope explicit).
ours_sig <- de %>%
  filter(!is.na(qvalue), qvalue <= 0.05,
         nutrient %in% c("C", "P")) %>%
  distinct(gene = `Gene names`, pSite, leading = `Leading proteins`)

# Residue letter + leading accession from the MaxQuant site table.
sty <- fread(file.path(repo_root, "data/Phospho (STY)Sites.csv"),
             select = c("Protein", "Gene names", "Amino acid", "Position")) %>%
  as_tibble() %>%
  mutate(pSite = paste(`Gene names`, Position, sep = "_")) %>%
  distinct(pSite, .keep_all = TRUE)

# Fallback for sites whose pSite key is degenerate (empty gene name -> "NA_524",
# multi-gene keys like "ENA2;ENA1_1065;1065"): join on leading accession +
# position instead of the string key.
sty_by_acc <- sty %>%
  transmute(acc_pos = paste(Protein, Position, sep = "_"),
            Protein2 = Protein, aa2 = `Amino acid`, pos2 = Position) %>%
  distinct(acc_pos, .keep_all = TRUE)

ours <- ours_sig %>%
  left_join(sty, by = "pSite") %>%
  mutate(acc_pos = paste(sub(";.*$", "", leading),
                         str_extract(pSite, "\\d+$"), sep = "_")) %>%
  left_join(sty_by_acc, by = "acc_pos") %>%
  mutate(Protein = coalesce(Protein, Protein2),
         position = as.integer(coalesce(Position, pos2)),
         aa = coalesce(`Amino acid`, aa2),
         orf = coalesce(uniprot_to_orf(Protein),
                        symbol_to_orf(gene)))
cat("This study: significant site records:", nrow(ours_sig),
    "| matched to STY table:", sum(!is.na(ours$aa)),
    "| mapped to ORF:", sum(!is.na(ours$orf)), "\n")

ours_sites <- ours %>%
  filter(!is.na(orf), !is.na(aa), !is.na(position)) %>%
  transmute(dataset = "ThisStudy", orf, gene, aa, position, pSite,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)

# ── Universe: all detected phosphosites (denominator for chi-square tests) ──
# Source: normed_pSite_ratio_perseus.csv contains every site that passed the
# localization probability threshold (> 0.7) in 05_psites_analysis.Rmd.
# We map each to an ORF+residue+position key using the same sty lookups above.
all_psites_raw <- read_csv(
  file.path(repo_root, "tables/normed_pSite_ratio_perseus.csv"),
  show_col_types = FALSE)

# Rename space-containing columns via base R before joining to sty (which also
# has a "Gene names" column); avoids .x/.y suffix collision and S4Vectors mask.
ap <- all_psites_raw %>%
  dplyr::select(pSite, Proteins, `Leading proteins`, `Gene names`)
names(ap)[names(ap) == "Gene names"]      <- "gene_name"
names(ap)[names(ap) == "Leading proteins"] <- "leading"

all_psites <- ap %>%
  left_join(sty, by = "pSite") %>%
  mutate(acc_pos  = paste(sub(";.*$", "", leading),
                          str_extract(pSite, "\\d+$"), sep = "_")) %>%
  left_join(sty_by_acc, by = "acc_pos") %>%
  mutate(Protein  = coalesce(Protein, Protein2),
         position = as.integer(coalesce(Position, pos2)),
         aa       = coalesce(`Amino acid`, aa2),
         orf      = coalesce(uniprot_to_orf(Protein),
                             symbol_to_orf(gene_name))) %>%
  filter(!is.na(orf), !is.na(aa), !is.na(position)) %>%
  transmute(key = site_key(orf, aa, position)) %>%
  distinct()

cat("Universe: all detected phosphosites with valid ORF+site mapping:",
    nrow(all_psites), "\n")
cat("Significant (rim15Δ-regulated, C or P, q≤0.05):", nrow(ours_sites), "\n")
```

# Li et al. 2019

```{r li2019}
li_raw <- read_excel(file.path(pub_dir, "Phosphoproteome/li2019/mmc5.xlsx"),
                     sheet = "Phospho Regulation Network") %>%
  setNames(c("del_gene", "psite_key", "log2fc")) %>%
  filter(grepl("rim15", del_gene, ignore.case = TRUE))

# keys look like NCL1_pT665pS667 -> explode into (T,665), (S,667)
li_sites <- li_raw %>%
  mutate(gene = sub("_p[STY].*$", "", psite_key),
         sites = str_extract_all(psite_key, "p[STY]\\d+")) %>%
  unnest(sites) %>%
  mutate(aa = substr(sites, 2, 2),
         position = as.integer(sub("^p[STY]", "", sites)),
         orf = symbol_to_orf(gene)) %>%
  filter(!is.na(orf)) %>%
  transmute(dataset = "Li2019", orf, gene, aa, position,
            log2fc_Li2019 = log2fc,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)
cat("Li 2019 (∆rim15 regulated): rows:", nrow(li_raw),
    "-> unique sites:", nrow(li_sites), "\n")
```

# Dokladal et al. 2021

```{r dokladal2021}
dok_raw <- suppressMessages(
  read_excel(file.path(pub_dir, "Phosphoproteome/dokladal2021/mmc3.xlsx"), sheet = "Exclusive",
             col_names = FALSE, skip = 3))
# rim15∆ block (after skipping the 3 header rows the empty first column is
# dropped, shifting everything left by one): col 23 = Protein, col 24 =
# Position, col 27 = 15-mer motif
dok_sites <- tibble(gene = dok_raw[[23]],
                    position = suppressWarnings(as.integer(dok_raw[[24]])),
                    motif = dok_raw[[27]]) %>%
  filter(!is.na(gene), !is.na(position)) %>%
  mutate(aa = toupper(substr(motif, 8, 8)),     # central residue of the 15-mer
         orf = symbol_to_orf(gene)) %>%
  filter(!is.na(orf), aa %in% c("S", "T", "Y")) %>%
  transmute(dataset = "Dokladal2021", orf, gene = toupper(gene), aa, position,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)
cat("Dokladal 2021 (Rim15 high-confidence targets): unique sites:",
    nrow(dok_sites), "\n")
```

# Baro et al. 2018

```{r baro2018}
baro_raw <- suppressMessages(
  read_excel(file.path(pub_dir, "Phosphoproteome/baro2018/Additional file 3.xlsx"),
             sheet = "Hoja1", col_names = FALSE, skip = 2)) %>%
  setNames(paste0("c", seq_len(ncol(.))))
# c1 Sequence | c2 Modifications | c7-c9 accessions (TiSH/SIMAC/TiO2) |
# c22-c24 gene symbols
baro_pep <- baro_raw %>%
  transmute(peptide = toupper(c1), mods = c2,
            accession = coalesce(c7, c8, c9),
            gene = coalesce(c22, c23, c24)) %>%
  filter(!is.na(peptide), !is.na(mods), !is.na(accession)) %>%
  mutate(accession = sub(";.*$", "", accession),
         start = locate_peptide(peptide, accession))
cat("Baro 2018 peptides:", nrow(baro_pep),
    "| located in UniProt sequence:", sum(!is.na(baro_pep$start)), "\n")

baro_sites <- baro_pep %>%
  filter(!is.na(start)) %>%
  mutate(site = str_extract_all(mods, "[STY]\\d+\\(Phospho\\)")) %>%
  unnest(site) %>%
  mutate(aa = substr(site, 1, 1),
         within = as.integer(str_extract(site, "\\d+")),
         position = start + within - 1L,
         orf = coalesce(uniprot_to_orf(accession), symbol_to_orf(gene))) %>%
  filter(!is.na(orf)) %>%
  # sanity: the residue letter must match the protein sequence at that position
  mutate(seq_aa = substr(proteome_chr[sub("-\\d+$", "", accession)],
                         position, position)) %>%
  filter(aa == seq_aa) %>%
  transmute(dataset = "Baro2018", orf, gene, aa, position,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)
cat("Baro 2018 (PP2A-Cdc55 hyperphosphorylated): unique sites:",
    nrow(baro_sites), "\n")
```

# Plank et al. (Cdc55-AID)

```{r plank}
plank_raw <- suppressWarnings(
  read_excel(file.path(pub_dir, "Phosphoproteome/plank2020/pmic13349-sup-0002-tables1.xlsx")))
plank_sig <- plank_raw %>%
  filter(is.na(Reverse), is.na(Potential.contaminant),
         Phospho..STY. >= 1,
         !is.na(pVal_aux_to_mock), pVal_aux_to_mock <= 0.05,
         meanlog2_aux_to_mock > 0) %>%
  transmute(peptide = toupper(Sequence),
            n_phos = as.integer(Phospho..STY.),
            accession = sub("^..\\|([^|]+)\\|.*$", "\\1",
                            sub(";.*$", "", Proteins)),
            log2fc_Plank = meanlog2_aux_to_mock,
            pval_Plank   = pVal_aux_to_mock)
cat("Plank Cdc55-AID: peptides nominal p<=0.05 & up on depletion:",
    nrow(plank_sig), "(NB: 1 peptide at pAdj<=0.05)\n")

plank_pep <- plank_sig %>%
  mutate(start = locate_peptide(peptide, accession),
         orf = uniprot_to_orf(accession),
         n_sty = str_count(peptide, "[STY]"))

# exact sites only where every S/T/Y in the peptide must be phosphorylated
plank_sites <- plank_pep %>%
  filter(!is.na(start), !is.na(orf), n_sty == n_phos) %>%
  mutate(within = map(peptide, ~ str_locate_all(.x, "[STY]")[[1]][, "start"])) %>%
  unnest_longer(within) %>%
  mutate(position = start + within - 1L,
         aa = substr(peptide, within, within)) %>%
  transmute(dataset = "Plank_Cdc55AID", orf,
            gene = NA_character_, aa, position,
            log2fc_Plank, pval_Plank,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)

plank_proteins <- plank_pep %>% filter(!is.na(orf)) %>% distinct(orf)
cat("Plank Cdc55-AID: unambiguous exact sites:", nrow(plank_sites),
    "| phosphoproteins (protein-level set):", nrow(plank_proteins), "\n")
```

# Overlap

```{r overlap}
all_sites <- bind_rows(ours_sites, li_sites, dok_sites, baro_sites, plank_sites)
set_names_ordered <- c("ThisStudy", "Li2019", "Dokladal2021",
                       "Baro2018", "Plank_Cdc55AID")

# guard: an empty dataset would otherwise silently drop its column
add_missing_sets <- function(df) {
  for (s in setdiff(set_names_ordered, names(df))) df[[s]] <- 0L
  df
}

# gene symbol per ORF for readability (prefer our own annotation)
orf_gene <- all_sites %>% filter(!is.na(gene)) %>% distinct(orf, gene) %>%
  group_by(orf) %>% dplyr::slice(1) %>% ungroup()

# --- exact-site-level membership ---
site_membership <- all_sites %>%
  distinct(dataset, key, orf, aa, position) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  add_missing_sets() %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(key, orf, gene, aa, position) %>%
  mutate(n_published = Li2019 + Dokladal2021 + Baro2018 + Plank_Cdc55AID,
         category = case_when(
           ThisStudy == 1 & n_published >= 2 ~ "consensus",
           ThisStudy == 1 & n_published >= 1 ~ "shared",
           ThisStudy == 1                    ~ "novel_this_study",
           TRUE                              ~ "published_only")) %>%
  arrange(desc(ThisStudy), desc(n_published), orf, position)
write_csv(site_membership, file.path(out_dir, "phosphosite_membership.csv"))

# --- protein-level membership (Plank contributes all phosphoproteins) ---
protein_sets <- bind_rows(
  all_sites %>% filter(dataset != "Plank_Cdc55AID") %>% distinct(dataset, orf),
  plank_proteins %>% mutate(dataset = "Plank_Cdc55AID"))
protein_membership <- protein_sets %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  add_missing_sets() %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  mutate(n_published = Li2019 + Dokladal2021 + Baro2018 + Plank_Cdc55AID) %>%
  arrange(desc(ThisStudy), desc(n_published), orf)
write_csv(protein_membership, file.path(out_dir, "phosphoprotein_membership.csv"))

pdf(file.path(out_dir, "phosphosite_overlap_upset.pdf"), width = 8, height = 5)
print(upset(as.data.frame(site_membership), sets = set_names_ordered,
            keep.order = TRUE, order.by = "freq",
            mainbar.y.label = "Exact phosphosites (ORF + residue)",
            sets.x.label = "Sites per dataset"))
dev.off()

pdf(file.path(out_dir, "phosphoprotein_overlap_upset.pdf"), width = 8, height = 5)
print(upset(as.data.frame(protein_membership), sets = set_names_ordered,
            keep.order = TRUE, order.by = "freq",
            mainbar.y.label = "Phosphoproteins (ORF)",
            sets.x.label = "Proteins per dataset"))
dev.off()

upset(as.data.frame(site_membership), sets = set_names_ordered,
      keep.order = TRUE, order.by = "freq",
      mainbar.y.label = "Exact phosphosites (ORF + residue)")
upset(as.data.frame(protein_membership), sets = set_names_ordered,
      keep.order = TRUE, order.by = "freq",
      mainbar.y.label = "Phosphoproteins (ORF)")
```

# Sanity check: Igo1 S64 and headline numbers

```{r headline}
# Igo1 = YNL157W; S64 is the canonical Rim15 site (Talarek et al. 2010)
site_membership %>% filter(orf == "YNL157W") %>% print(n = 50)

counts <- site_membership %>% summarise(
  ours_total          = sum(ThisStudy),
  consensus_ge2       = sum(category == "consensus"),
  shared_ge1          = sum(category %in% c("consensus", "shared")),
  novel               = sum(category == "novel_this_study"))
print(counts)

cat("\nProtein-level: our phosphoproteins:",
    sum(protein_membership$ThisStudy == 1),
    "| overlapping >=1 published:",
    sum(protein_membership$ThisStudy == 1 & protein_membership$n_published >= 1),
    "| >=2 published:",
    sum(protein_membership$ThisStudy == 1 & protein_membership$n_published >= 2),
    "\n")

# our sites shared with each published dataset (site-level)
for (s in setdiff(set_names_ordered, "ThisStudy"))
  cat("ThisStudy ∩", s, ":",
      sum(site_membership$ThisStudy == 1 & site_membership[[s]] == 1), "\n")
```

# Statistical tests

```{r statistics}
# Chi-square test for each published dataset.
# Universe = all phosphosites detected in our study with valid ORF+site mapping.
# Question: are Rim15-regulated sites (q≤0.05 in C or P starvation) enriched
# among sites that appear in each published Rim15 / PP2A-Cdc55 dataset?
#
# Contingency table for each published set X:
#   [ sig ∩ X      sig ∩ not-X ]
#   [ not-sig ∩ X  not-sig ∩ not-X ]

published_sets <- list(
  Li2019        = li_sites$key,
  Dokladal2021  = dok_sites$key,
  Baro2018      = baro_sites$key,
  Plank_Cdc55AID = plank_sites$key
)

universe_keys <- all_psites$key
sig_keys      <- ours_sites$key

chisq_results <- map_dfr(names(published_sets), function(nm) {
  pub_keys <- published_sets[[nm]]
  a <- sum(universe_keys %in% sig_keys &  universe_keys %in% pub_keys)
  b <- sum(universe_keys %in% sig_keys & !universe_keys %in% pub_keys)
  c <- sum(!universe_keys %in% sig_keys &  universe_keys %in% pub_keys)
  d <- sum(!universe_keys %in% sig_keys & !universe_keys %in% pub_keys)
  ct <- chisq.test(matrix(c(a, c, b, d), nrow = 2))
  tibble(
    dataset   = nm,
    n_published = a + c,
    n_sig     = a + b,
    n_overlap = a,
    N         = length(universe_keys),
    pct_sig_in_pub = round(100 * a / (a + b), 1),
    pct_nonsig_in_pub = round(100 * c / (c + d), 1),
    chi_sq    = round(ct$statistic, 2),
    p_value   = ct$p.value
  )
})

print(chisq_results)
```

# Quantitative overlap table

```{r quantitative_overlap}
# Phosphosites in our significant set (q≤0.05, C or P) that appear in ≥1
# published dataset, with fold-change from each study side by side.
#
# Our metrics: best (lowest-q) contrast per site from the DE table.
# de columns: pSite, nutrient, time, pvalue, qvalue, log2FC
de_best <- de %>%
  filter(!is.na(qvalue)) %>%
  group_by(pSite) %>%
  slice_min(qvalue, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(pSite,
            log2FC_thisStudy = log2FC,
            qvalue_thisStudy = qvalue,
            nutrient_thisStudy = nutrient,
            time_thisStudy = time)

# ours_sites carries pSite; join to get our quantitative values
ours_fc <- ours_sites %>%
  left_join(de_best, by = "pSite")

# Lookup tables: quantitative metrics from published datasets (keyed on site_key)
li_fc_lookup    <- li_sites    %>% transmute(key, log2fc_Li2019)
plank_fc_lookup <- plank_sites %>% transmute(key, log2fc_Plank, pval_Plank)

# Build the overlap table: our sig sites ∩ ≥1 published dataset
overlap_psites <- site_membership %>%
  filter(ThisStudy == 1, n_published >= 1) %>%
  left_join(ours_fc %>%
              transmute(key, log2FC_thisStudy, qvalue_thisStudy,
                        nutrient_thisStudy, time_thisStudy),
            by = "key") %>%
  left_join(li_fc_lookup,    by = "key") %>%
  left_join(plank_fc_lookup, by = "key") %>%
  dplyr::select(key, orf, gene, aa, position,
                log2FC_thisStudy, qvalue_thisStudy, nutrient_thisStudy, time_thisStudy,
                Li2019, log2fc_Li2019,
                Dokladal2021, Baro2018,
                Plank_Cdc55AID, log2fc_Plank, pval_Plank,
                n_published, category) %>%
  arrange(desc(n_published), qvalue_thisStudy)

write_csv(overlap_psites, file.path(out_dir, "phosphosite_overlap_quantitative.csv"))
cat("Quantitative overlap table:", nrow(overlap_psites),
    "sites (ThisStudy ∩ ≥1 published dataset)\n")
cat("  Of these, log2fc_Li2019 available for:",
    sum(!is.na(overlap_psites$log2fc_Li2019)), "sites\n")
cat("  log2fc_Plank available for:",
    sum(!is.na(overlap_psites$log2fc_Plank)), "sites\n")
```

# Summary paragraph

```{r paragraph, results='asis'}
fmt_p <- function(p) {
  if (p < 0.001) formatC(p, format = "e", digits = 1) else round(p, 3)
}

n_universe <- nrow(all_psites)
n_sig      <- nrow(ours_sites)

# Collect key numbers for each dataset
r  <- chisq_results

r_li  <- r[r$dataset == "Li2019",]
r_dok <- r[r$dataset == "Dokladal2021",]
r_bar <- r[r$dataset == "Baro2018",]
r_plk <- r[r$dataset == "Plank_Cdc55AID",]

cat(sprintf(
"To evaluate whether the rim15Δ-regulated phosphosites identified in this
study are consistent with previously reported Rim15 targets, we compared our
significant sites (q ≤ 0.05 in carbon or phosphorus starvation; n = %d) with
four published phosphoproteomic datasets, using a chi-square test with all
%d phosphosites detected in our experiment as the universe. Our sites were
significantly enriched among those identified by Dokladal et al. (2021;
%d/%d overlap, χ² = %.1f, p = %s) and Baro et al. (2018; %d/%d overlap,
χ² = %.1f, p = %s). Overlap with Li et al. (2019; %d/%d sites, χ² = %.1f,
p = %s) and Plank et al. (2020; %d/%d sites, χ² = %.1f, p = %s) was not
statistically significant. The significant enrichment with two independent
datasets using distinct experimental strategies supports the validity of our
phosphoproteomic approach for identifying Rim15-dependent phosphorylation events.",
  n_sig, n_universe,
  r_dok$n_overlap, r_dok$n_sig, r_dok$chi_sq, fmt_p(r_dok$p_value),
  r_bar$n_overlap, r_bar$n_sig, r_bar$chi_sq, fmt_p(r_bar$p_value),
  r_li$n_overlap,  r_li$n_sig,  r_li$chi_sq,  fmt_p(r_li$p_value),
  r_plk$n_overlap, r_plk$n_sig, r_plk$chi_sq, fmt_p(r_plk$p_value))))
```

```{r session}
sessionInfo()
```
