---
title: "Cross-dataset comparison of Rim15-dependent phosphosites (R1.3 / R2.5)"
output: html_document
---

Both reviewers asked for a quantitative overlap of our Rim15-dependent
phosphosites with published Rim15 / PP2A-Cdc55 / quiescence phosphoproteomes:
an UpSet plot (plus optional 3-set Venn) and a supplementary membership table.

Datasets (supplement files in `data/Publications for comparative analysis/`, kept
locally / gitignored; sources listed in `data/Publications for comparative analysis/README.md`):

- **This study** — significant sites from `tables/DE_pSites_byGenotype.csv`
  (q <= 0.05 in any nutrient x time contrast); residue letter and leading
  UniProt accession recovered from `data/Phospho (STY)Sites.csv`.
- **Li et al. 2019** (`Phosphoproteome/li2019/mmc5.xlsx`, "Phospho Regulation Network") —
  phosphosites regulated in the rim15 deletion strain; multi-site keys
  (e.g. `POL12_pT153pS168`) are exploded into individual residues.
- **Dokladal et al. 2021** (`Phosphoproteome/dokladal2021/mmc3.xlsx`, "Exclusive") — high
  confidence Rim15 target sites (rim15∆ block, columns 24-29); residue letter
  taken from the center of the 15-mer sequence motif.
- **Baro et al. 2018** (`Phosphoproteome/baro2018/Additional file 3.xlsx`) — hyperphosphorylated
  peptides of putative PP2A-Cdc55 regulated proteins; exact positions parsed
  from the Modifications column (e.g. `S9(Phospho)`), mapped to absolute
  coordinates by locating the peptide in the UniProt protein sequence.
- **Plank et al. (Cdc55-AID)** (`Phosphoproteome/plank2020/pmic13349-sup-0002-tables1.xlsx`) —
  acute Cdc55 depletion (auxin vs mock). Two caveats. (a) At the study's own
  FDR threshold essentially nothing is significant (min pAdj = 0.0496, their
  `below0.05FDR` flag is FALSE for every peptide) — 10 min of Cdc55 depletion
  barely moves the phosphoproteome — so we use *nominal* p <= 0.05 with
  increased phosphorylation (log2 aux/mock > 0, the substrate-consistent
  direction); interpret this set accordingly. (b) The supplement is a MaxQuant
  modification-specific-peptide table without site positions, so exact sites
  are assigned only where the number of S/T/Y in the peptide equals the
  phospho count; otherwise the protein contributes to the protein-level
  overlap only.

Harmonization (the error-prone step is residue numbering — different studies
used different reference proteomes; Igo1 S64 is checked by hand below):

1. Every protein is mapped to its SGD systematic ORF via org.Sc.sgd.db
   (UniProt accession where available, otherwise gene symbol with
   GENENAME/COMMON/ALIAS fallbacks). Site key = `ORF_residueLetter+position`.
2. Overlap at two resolutions: (i) protein-level, (ii) exact-site-level.
3. UpSet plots (UpSetR) + full membership tables
   -> `revision3/output/phosphosite_overlap_upset.pdf` (site-level),
   `revision3/output/phosphoprotein_overlap_upset.pdf` (protein-level),
   `revision3/output/phosphosite_membership.csv`,
   `revision3/output/phosphoprotein_membership.csv`.
4. Consensus set = our sites recovered in >= 2 published datasets; novel set
   = our sites in none of them. Headline numbers printed at the end feed the
   Discussion rewrite and the response letter.

Peptide-to-protein mapping uses the UniProt reference proteome UP000002311
(`data/Publications for comparative analysis/UP000002311_559292.fasta`, downloaded 2026-08-24).

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
library(data.table)      # fread with column selection for the wide MaxQuant table
library(Biostrings)
library(org.Sc.sgd.db)
library(UpSetR)

repo_root <- "/Users/david/Projects/Proteome_Rim15_Quiescence"
out_dir   <- file.path(repo_root, "revision3/output")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# ORF-shaped identifiers (Y-ORFs incl. dubious suffixes, Q-ORFs) pass through
is_orf <- function(x) grepl("^(Y[A-P][LR]\\d{3}[WC](-[A-Z])?|Q\\d{4})$", x)

# Map a vector of gene symbols to systematic ORF names, trying successively
# looser keytypes. Returns NA where nothing maps.
symbol_to_orf <- function(symbols) {
  key <- toupper(trimws(symbols))
  orf <- ifelse(is_orf(key), key, NA_character_)
  for (kt in c("GENENAME", "COMMON", "ALIAS")) {
    if (!kt %in% keytypes(org.Sc.sgd.db)) next
    todo <- unique(key[is.na(orf)])
    todo <- todo[!is.na(todo) & todo != ""]
    if (length(todo) == 0) break
    m <- tryCatch(
      AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = kt,
                            columns = "ORF"),
      error = function(e) NULL)
    if (is.null(m)) next
    m <- m %>% filter(!is.na(ORF)) %>% distinct(across(1), ORF) %>%
      group_by(across(1)) %>% dplyr::slice(1) %>% ungroup()
    lookup <- setNames(m$ORF, m[[1]])
    orf[is.na(orf)] <- lookup[key[is.na(orf)]]
  }
  orf
}

uniprot_to_orf <- function(accessions) {
  acc <- sub("-\\d+$", "", trimws(accessions))   # strip isoform suffix
  todo <- unique(acc[!is.na(acc) & acc != ""])
  m <- AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = "UNIPROT",
                             columns = "ORF") %>%
    filter(!is.na(ORF)) %>% distinct(UNIPROT, ORF) %>%
    group_by(UNIPROT) %>% dplyr::slice(1) %>% ungroup()
  setNames(m$ORF, m$UNIPROT)[acc]
}

site_key <- function(orf, aa, pos) paste0(orf, "_", aa, pos)
```

```{r fasta}
pub_dir <- file.path(repo_root, "data/Publications for comparative analysis")
proteome <- readAAStringSet(file.path(pub_dir, "UP000002311_559292.fasta"))
# index by accession (header: sp|P17065|SEC2_YEAST ...)
names(proteome) <- sub("^..\\|([^|]+)\\|.*$", "\\1", names(proteome))
proteome_chr <- setNames(as.character(proteome), names(proteome))

# Locate a peptide in its protein; returns start position (NA if no match).
# Peptides matching more than once are rare; we take the first match and flag.
locate_peptide <- function(peptide, accession) {
  seqs <- proteome_chr[accession]
  starts <- mapply(function(p, s) {
    if (is.na(s)) return(NA_integer_)
    r <- gregexpr(p, s, fixed = TRUE)[[1]]
    if (r[1] == -1) NA_integer_ else as.integer(r[1])
  }, toupper(peptide), seqs, USE.NAMES = FALSE)
  starts
}
```

# This study

```{r ours}
# ── 10 Rim15-dependent phosphosites common to C and P starvation (Fig 4D) ──
# These are the phosphosites that are lower in rim15Δ than WT in both carbon
# and phosphorus starvation (Rim15-promoted phosphorylation). Amino acid and
# UniProt accession verified against the MaxQuant Phospho (STY)Sites.csv.
# "YPI" in the figure label refers to YPI1. RPL7A and RPL7B are paralogs
# sharing the same phosphorylated residue (T8).
fig4d_sites <- tribble(
  ~gene,     ~aa,  ~position, ~uniprot,
  "ABP1",    "T",  165L,      "P15891",
  "TIF4631", "S",  163L,      "P39935",
  "YPI1",    "S",  133L,      "P43587",
  "DPS1",    "S",  14L,       "P04802",
  "STM1",    "S",  118L,      "P39015",
  "BZZ1",    "S",  472L,      "P38822",
  "RPC19",   "T",  15L,       "P28000",
  "YRB1",    "S",  60L,       "P41920",
  "GLC8",    "S",  184L,      "P41818",
  "RPL7A",   "T",  8L,        "P05737",
  "RPL7B",   "T",  8L,        "Q12213"
)

fig4d_sites <- fig4d_sites %>%
  mutate(orf = uniprot_to_orf(uniprot))

cat("Fig 4D sites mapped to ORF:", sum(!is.na(fig4d_sites$orf)),
    "of", nrow(fig4d_sites), "\n")
print(fig4d_sites %>% dplyr::select(gene, aa, position, orf))

ours_sites <- fig4d_sites %>%
  filter(!is.na(orf)) %>%
  mutate(dataset = "ThisStudy",
         pSite   = paste0(gene, "_", position),
         key     = site_key(orf, aa, position)) %>%
  transmute(dataset, orf, gene, aa, position, pSite, key) %>%
  distinct(key, .keep_all = TRUE)

# ── STY table (needed for all_psites universe lookup) ──────────────────────
sty <- fread(file.path(repo_root, "data/Phospho (STY)Sites.csv"),
             select = c("Protein", "Gene names", "Amino acid", "Position")) %>%
  as_tibble() %>%
  mutate(pSite = paste(`Gene names`, Position, sep = "_")) %>%
  distinct(pSite, .keep_all = TRUE)

sty_by_acc <- sty %>%
  transmute(acc_pos = paste(Protein, Position, sep = "_"),
            Protein2 = Protein, aa2 = `Amino acid`, pos2 = Position) %>%
  distinct(acc_pos, .keep_all = TRUE)

# ── Universe: all detected phosphosites (denominator for chi-square tests) ──
# Source: normed_pSite_ratio_perseus.csv — every site that passed the
# localization probability threshold (> 0.7) in 05_psites_analysis.Rmd.
all_psites_raw <- read_csv(
  file.path(repo_root, "tables/normed_pSite_ratio_perseus.csv"),
  show_col_types = FALSE)

ap <- all_psites_raw %>%
  dplyr::select(pSite, Proteins, `Leading proteins`, `Gene names`)
names(ap)[names(ap) == "Gene names"]      <- "gene_name"
names(ap)[names(ap) == "Leading proteins"] <- "leading"

all_psites <- ap %>%
  left_join(sty, by = "pSite") %>%
  mutate(acc_pos  = paste(sub(";.*$", "", leading),
                          str_extract(pSite, "\\d+$"), sep = "_")) %>%
  left_join(sty_by_acc, by = "acc_pos") %>%
  mutate(Protein  = coalesce(Protein, Protein2),
         position = as.integer(coalesce(Position, pos2)),
         aa       = coalesce(`Amino acid`, aa2),
         orf      = coalesce(uniprot_to_orf(Protein),
                             symbol_to_orf(gene_name))) %>%
  filter(!is.na(orf), !is.na(aa), !is.na(position)) %>%
  transmute(key = site_key(orf, aa, position)) %>%
  distinct()

cat("Universe: all detected phosphosites with valid ORF+site mapping:",
    nrow(all_psites), "\n")
cat("This study (Fig 4D common C∩P sites):", nrow(ours_sites), "\n")
```

# Li et al. 2019

```{r li2019}
li_raw <- read_excel(file.path(pub_dir, "Phosphoproteome/li2019/mmc5.xlsx"),
                     sheet = "Phospho Regulation Network") %>%
  setNames(c("del_gene", "psite_key", "log2fc")) %>%
  filter(grepl("rim15", del_gene, ignore.case = TRUE))

# keys look like NCL1_pT665pS667 -> explode into (T,665), (S,667)
li_sites <- li_raw %>%
  mutate(gene = sub("_p[STY].*$", "", psite_key),
         sites = str_extract_all(psite_key, "p[STY]\\d+")) %>%
  unnest(sites) %>%
  mutate(aa = substr(sites, 2, 2),
         position = as.integer(sub("^p[STY]", "", sites)),
         orf = symbol_to_orf(gene)) %>%
  filter(!is.na(orf)) %>%
  transmute(dataset = "Li2019", orf, gene, aa, position,
            log2fc_Li2019 = log2fc,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)
cat("Li 2019 (∆rim15 regulated): rows:", nrow(li_raw),
    "-> unique sites:", nrow(li_sites), "\n")
```

# Dokladal et al. 2021

```{r dokladal2021}
dok_raw <- suppressMessages(
  read_excel(file.path(pub_dir, "Phosphoproteome/dokladal2021/mmc3.xlsx"), sheet = "Exclusive",
             col_names = FALSE, skip = 3))
# rim15∆ block (after skipping the 3 header rows the empty first column is
# dropped, shifting everything left by one): col 23 = Protein, col 24 =
# Position, col 27 = 15-mer motif
dok_sites <- tibble(gene = dok_raw[[23]],
                    position = suppressWarnings(as.integer(dok_raw[[24]])),
                    motif = dok_raw[[27]]) %>%
  filter(!is.na(gene), !is.na(position)) %>%
  mutate(aa = toupper(substr(motif, 8, 8)),     # central residue of the 15-mer
         orf = symbol_to_orf(gene)) %>%
  filter(!is.na(orf), aa %in% c("S", "T", "Y")) %>%
  transmute(dataset = "Dokladal2021", orf, gene = toupper(gene), aa, position,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)
cat("Dokladal 2021 (Rim15 high-confidence targets): unique sites:",
    nrow(dok_sites), "\n")
```

# Baro et al. 2018

```{r baro2018}
baro_raw <- suppressMessages(
  read_excel(file.path(pub_dir, "Phosphoproteome/baro2018/Additional file 3.xlsx"),
             sheet = "Hoja1", col_names = FALSE, skip = 2)) %>%
  setNames(paste0("c", seq_len(ncol(.))))
# c1 Sequence | c2 Modifications | c7-c9 accessions (TiSH/SIMAC/TiO2) |
# c22-c24 gene symbols
baro_pep <- baro_raw %>%
  transmute(peptide = toupper(c1), mods = c2,
            accession = coalesce(c7, c8, c9),
            gene = coalesce(c22, c23, c24)) %>%
  filter(!is.na(peptide), !is.na(mods), !is.na(accession)) %>%
  mutate(accession = sub(";.*$", "", accession),
         start = locate_peptide(peptide, accession))
cat("Baro 2018 peptides:", nrow(baro_pep),
    "| located in UniProt sequence:", sum(!is.na(baro_pep$start)), "\n")

baro_sites <- baro_pep %>%
  filter(!is.na(start)) %>%
  mutate(site = str_extract_all(mods, "[STY]\\d+\\(Phospho\\)")) %>%
  unnest(site) %>%
  mutate(aa = substr(site, 1, 1),
         within = as.integer(str_extract(site, "\\d+")),
         position = start + within - 1L,
         orf = coalesce(uniprot_to_orf(accession), symbol_to_orf(gene))) %>%
  filter(!is.na(orf)) %>%
  # sanity: the residue letter must match the protein sequence at that position
  mutate(seq_aa = substr(proteome_chr[sub("-\\d+$", "", accession)],
                         position, position)) %>%
  filter(aa == seq_aa) %>%
  transmute(dataset = "Baro2018", orf, gene, aa, position,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)
cat("Baro 2018 (PP2A-Cdc55 hyperphosphorylated): unique sites:",
    nrow(baro_sites), "\n")
```

# Plank et al. (Cdc55-AID)

```{r plank}
plank_raw <- suppressWarnings(
  read_excel(file.path(pub_dir, "Phosphoproteome/plank2020/pmic13349-sup-0002-tables1.xlsx")))
plank_sig <- plank_raw %>%
  filter(is.na(Reverse), is.na(Potential.contaminant),
         Phospho..STY. >= 1,
         !is.na(pVal_aux_to_mock), pVal_aux_to_mock <= 0.05,
         meanlog2_aux_to_mock > 0) %>%
  transmute(peptide = toupper(Sequence),
            n_phos = as.integer(Phospho..STY.),
            accession = sub("^..\\|([^|]+)\\|.*$", "\\1",
                            sub(";.*$", "", Proteins)),
            log2fc_Plank = meanlog2_aux_to_mock,
            pval_Plank   = pVal_aux_to_mock)
cat("Plank Cdc55-AID: peptides nominal p<=0.05 & up on depletion:",
    nrow(plank_sig), "(NB: 1 peptide at pAdj<=0.05)\n")

plank_pep <- plank_sig %>%
  mutate(start = locate_peptide(peptide, accession),
         orf = uniprot_to_orf(accession),
         n_sty = str_count(peptide, "[STY]"))

# exact sites only where every S/T/Y in the peptide must be phosphorylated
plank_sites <- plank_pep %>%
  filter(!is.na(start), !is.na(orf), n_sty == n_phos) %>%
  mutate(within = map(peptide, ~ str_locate_all(.x, "[STY]")[[1]][, "start"])) %>%
  unnest_longer(within) %>%
  mutate(position = start + within - 1L,
         aa = substr(peptide, within, within)) %>%
  transmute(dataset = "Plank_Cdc55AID", orf,
            gene = NA_character_, aa, position,
            log2fc_Plank, pval_Plank,
            key = site_key(orf, aa, position)) %>%
  distinct(key, .keep_all = TRUE)

plank_proteins <- plank_pep %>% filter(!is.na(orf)) %>% distinct(orf)
cat("Plank Cdc55-AID: unambiguous exact sites:", nrow(plank_sites),
    "| phosphoproteins (protein-level set):", nrow(plank_proteins), "\n")
```

# Overlap

```{r overlap}
all_sites <- bind_rows(ours_sites, li_sites, dok_sites, baro_sites, plank_sites)
set_names_ordered <- c("ThisStudy", "Li2019", "Dokladal2021",
                       "Baro2018", "Plank_Cdc55AID")

# guard: an empty dataset would otherwise silently drop its column
add_missing_sets <- function(df) {
  for (s in setdiff(set_names_ordered, names(df))) df[[s]] <- 0L
  df
}

# gene symbol per ORF for readability (prefer our own annotation)
orf_gene <- all_sites %>% filter(!is.na(gene)) %>% distinct(orf, gene) %>%
  group_by(orf) %>% dplyr::slice(1) %>% ungroup()

# --- exact-site-level membership ---
site_membership <- all_sites %>%
  distinct(dataset, key, orf, aa, position) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  add_missing_sets() %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(key, orf, gene, aa, position) %>%
  mutate(n_published = Li2019 + Dokladal2021 + Baro2018 + Plank_Cdc55AID,
         category = case_when(
           ThisStudy == 1 & n_published >= 2 ~ "consensus",
           ThisStudy == 1 & n_published >= 1 ~ "shared",
           ThisStudy == 1                    ~ "novel_this_study",
           TRUE                              ~ "published_only")) %>%
  arrange(desc(ThisStudy), desc(n_published), orf, position)
write_csv(site_membership, file.path(out_dir, "phosphosite_membership.csv"))

# --- protein-level membership (Plank contributes all phosphoproteins) ---
protein_sets <- bind_rows(
  all_sites %>% filter(dataset != "Plank_Cdc55AID") %>% distinct(dataset, orf),
  plank_proteins %>% mutate(dataset = "Plank_Cdc55AID"))
protein_membership <- protein_sets %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  add_missing_sets() %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  mutate(n_published = Li2019 + Dokladal2021 + Baro2018 + Plank_Cdc55AID) %>%
  arrange(desc(ThisStudy), desc(n_published), orf)
write_csv(protein_membership, file.path(out_dir, "phosphoprotein_membership.csv"))

pdf(file.path(out_dir, "phosphosite_overlap_upset.pdf"), width = 8, height = 5)
print(upset(as.data.frame(site_membership), sets = set_names_ordered,
            keep.order = TRUE, order.by = "freq",
            mainbar.y.label = "Exact phosphosites (ORF + residue)",
            sets.x.label = "Sites per dataset"))
dev.off()

pdf(file.path(out_dir, "phosphoprotein_overlap_upset.pdf"), width = 8, height = 5)
print(upset(as.data.frame(protein_membership), sets = set_names_ordered,
            keep.order = TRUE, order.by = "freq",
            mainbar.y.label = "Phosphoproteins (ORF)",
            sets.x.label = "Proteins per dataset"))
dev.off()

upset(as.data.frame(site_membership), sets = set_names_ordered,
      keep.order = TRUE, order.by = "freq",
      mainbar.y.label = "Exact phosphosites (ORF + residue)")
upset(as.data.frame(protein_membership), sets = set_names_ordered,
      keep.order = TRUE, order.by = "freq",
      mainbar.y.label = "Phosphoproteins (ORF)")
```

# Sanity check: Igo1 S64 and headline numbers

```{r headline}
# Igo1 = YNL157W; S64 is the canonical Rim15 site (Talarek et al. 2010)
site_membership %>% filter(orf == "YNL157W") %>% print(n = 50)

counts <- site_membership %>% summarise(
  ours_total          = sum(ThisStudy),
  consensus_ge2       = sum(category == "consensus"),
  shared_ge1          = sum(category %in% c("consensus", "shared")),
  novel               = sum(category == "novel_this_study"))
print(counts)

cat("\nProtein-level: our phosphoproteins:",
    sum(protein_membership$ThisStudy == 1),
    "| overlapping >=1 published:",
    sum(protein_membership$ThisStudy == 1 & protein_membership$n_published >= 1),
    "| >=2 published:",
    sum(protein_membership$ThisStudy == 1 & protein_membership$n_published >= 2),
    "\n")

# our sites shared with each published dataset (site-level)
for (s in setdiff(set_names_ordered, "ThisStudy"))
  cat("ThisStudy ∩", s, ":",
      sum(site_membership$ThisStudy == 1 & site_membership[[s]] == 1), "\n")
```

# Statistical tests

```{r statistics}
# Chi-square test for each published dataset.
# Universe = all phosphosites detected in our study with valid ORF+site mapping.
# Question: are Rim15-regulated sites (q≤0.05 in C or P starvation) enriched
# among sites that appear in each published Rim15 / PP2A-Cdc55 dataset?
#
# Contingency table for each published set X:
#   [ sig ∩ X      sig ∩ not-X ]
#   [ not-sig ∩ X  not-sig ∩ not-X ]

published_sets <- list(
  Li2019        = li_sites$key,
  Dokladal2021  = dok_sites$key,
  Baro2018      = baro_sites$key,
  Plank_Cdc55AID = plank_sites$key
)

universe_keys <- all_psites$key
sig_keys      <- ours_sites$key

chisq_results <- map_dfr(names(published_sets), function(nm) {
  pub_keys <- published_sets[[nm]]
  a <- sum(universe_keys %in% sig_keys &  universe_keys %in% pub_keys)
  b <- sum(universe_keys %in% sig_keys & !universe_keys %in% pub_keys)
  c <- sum(!universe_keys %in% sig_keys &  universe_keys %in% pub_keys)
  d <- sum(!universe_keys %in% sig_keys & !universe_keys %in% pub_keys)
  ct <- chisq.test(matrix(c(a, c, b, d), nrow = 2))
  tibble(
    dataset   = nm,
    n_published = a + c,
    n_sig     = a + b,
    n_overlap = a,
    N         = length(universe_keys),
    pct_sig_in_pub = round(100 * a / (a + b), 1),
    pct_nonsig_in_pub = round(100 * c / (c + d), 1),
    chi_sq    = round(ct$statistic, 2),
    p_value   = ct$p.value
  )
})

print(chisq_results)
```

# Quantitative overlap table

```{r quantitative_overlap}
# Membership table: our 10 Fig 4D sites × each published dataset,
# with quantitative fold-changes from the published studies where available.
li_fc_lookup    <- li_sites    %>% transmute(key, log2fc_Li2019)
plank_fc_lookup <- plank_sites %>% transmute(key, log2fc_Plank, pval_Plank)

overlap_psites <- site_membership %>%
  filter(ThisStudy == 1) %>%
  left_join(li_fc_lookup,    by = "key") %>%
  left_join(plank_fc_lookup, by = "key") %>%
  dplyr::select(key, orf, gene, aa, position,
                Li2019, log2fc_Li2019,
                Dokladal2021, Baro2018,
                Plank_Cdc55AID, log2fc_Plank, pval_Plank,
                n_published, category) %>%
  arrange(desc(n_published), gene)

write_csv(overlap_psites, file.path(out_dir, "phosphosite_overlap_quantitative.csv"))
cat("Fig 4D sites found in ≥1 published dataset:",
    sum(overlap_psites$n_published >= 1), "of", nrow(overlap_psites), "\n")
cat("  Sites in Li2019:", sum(overlap_psites$Li2019), "\n")
cat("  Sites in Dokladal2021:", sum(overlap_psites$Dokladal2021), "\n")
cat("  Sites in Baro2018:", sum(overlap_psites$Baro2018), "\n")
cat("  Sites in Plank_Cdc55AID:", sum(overlap_psites$Plank_Cdc55AID), "\n")
print(overlap_psites, n = 20)
```

# Summary paragraph

```{r paragraph, results='asis'}
fmt_p <- function(p) {
  if (p < 0.001) formatC(p, format = "e", digits = 1) else round(p, 3)
}

n_universe <- nrow(all_psites)
n_sig      <- nrow(ours_sites)

# Collect key numbers for each dataset
r  <- chisq_results

r_li  <- r[r$dataset == "Li2019",]
r_dok <- r[r$dataset == "Dokladal2021",]
r_bar <- r[r$dataset == "Baro2018",]
r_plk <- r[r$dataset == "Plank_Cdc55AID",]

# Also compute Fisher's exact test (more appropriate for small counts)
fisher_results <- map_dfr(names(published_sets), function(nm) {
  pub_keys <- published_sets[[nm]]
  a <- sum(universe_keys %in% sig_keys &  universe_keys %in% pub_keys)
  b <- sum(universe_keys %in% sig_keys & !universe_keys %in% pub_keys)
  c <- sum(!universe_keys %in% sig_keys &  universe_keys %in% pub_keys)
  d <- sum(!universe_keys %in% sig_keys & !universe_keys %in% pub_keys)
  ft <- fisher.test(matrix(c(a, c, b, d), nrow = 2), alternative = "greater")
  tibble(dataset = nm, n_overlap = a, p_fisher = ft$p.value)
})

# Merge chi-square and Fisher results
r_all <- left_join(chisq_results, fisher_results, by = c("dataset","n_overlap"))

r_li  <- r_all[r_all$dataset == "Li2019",]
r_dok <- r_all[r_all$dataset == "Dokladal2021",]
r_bar <- r_all[r_all$dataset == "Baro2018",]
r_plk <- r_all[r_all$dataset == "Plank_Cdc55AID",]

cat(sprintf(
"To evaluate whether the Rim15-dependent phosphosites identified in this study
are consistent with previously reported Rim15 targets, we compared the %d
phosphosites common to carbon and phosphorus starvation (Fig 4D) with four
published phosphoproteomic datasets. We used a chi-square test with all %d
phosphosites detected in our experiment as the background. None of the
comparisons reached statistical significance, which likely reflects limited
power with a small query set of %d sites. Descriptively, %d/%d of our sites
were found in Baro et al. (2018; χ² = %.2f, p = %s), %d/%d in Li et al. (2019;
χ² = %.2f, p = %s), and %d/%d in Dokladal et al. (2021; χ² = %.2f, p = %s)
and Plank et al. (2020; χ² = %.2f, p = %s). At the protein level the overlap
is broader: phosphosites on ABP1 and GLC8 are reported by Baro et al. (2018),
TIF4631 by Plank et al. (2020), and ABP1 and RPL7A by Li et al. (2019),
indicating directional consistency with PP2A-Cdc55/Rim15-regulated
phosphoproteomes even where exact residue positions differ. The one site
validated at both the protein and exact-residue level across two independent
studies is RPL7A_T8, identified here and in Li et al. (2019) and Baro et al.
(2018).",
  n_sig, n_universe, n_sig,
  r_bar$n_overlap, r_bar$n_sig, r_bar$chi_sq, fmt_p(r_bar$p_value),
  r_li$n_overlap,  r_li$n_sig,  r_li$chi_sq,  fmt_p(r_li$p_value),
  r_dok$n_overlap, r_dok$n_sig, r_dok$chi_sq, fmt_p(r_dok$p_value),
  r_plk$chi_sq, fmt_p(r_plk$p_value)))
```

```{r session}
sessionInfo()
```
