---
title: "Cross-dataset comparison of quiescence proteome (R2.3)"
output: html_document
---

Reviewer 2 requested a systematic comparison of the proteins we detected and
the proteins regulated in rim15Δ with those reported in published quiescence
proteome studies.

**Source tables from this study:**

- **Detected proteins** — `tables/supplemental/S5_normalized_protein_ratios.csv`
  (Table S5): all proteins with at least one quantified SILAC ratio across the
  experiment. UniProt leading accessions in the `Protein IDs` column.
- **rim15Δ-regulated proteins** — `tables/DE_proteins_byGenotype.csv`
  (Table S16): proteins significantly different between WT and rim15Δ
  (q ≤ 0.05 in any nutrient × timepoint contrast); log2FC > 0 = higher in WT.

**Published datasets:**

- **Paulo et al. 2015** (Mol Cell Proteomics 14:1998–2015,
  `Proteome/Paulo et al., 2015/`) — label-free quant of quiescent vs
  non-quiescent yeast (sedimentation density fractionation). Two outputs used:
  (a) `NIHMS464126-supplement-2_si_001.xlsx` Sheet1 — all 2438 detected ORFs;
  (b) `NIHMS464126-supplement-3_si_003.xlsx` sheet "2X or more Up in Quiescent"
  — proteins significantly and ≥2-fold enriched in quiescent cells.
- **den Ridder et al. 2023** (Biotechnol Biofuels 16, `Proteome/den Ridder et al., 2023/`)
  — TMT-based proteome of CEN.PK113-7D and IMX372 yeast under aerobic and
  anaerobic conditions (not a quiescence study; included for proteome coverage
  comparison only). `mmc7.xlsx` "All proteins" — 1739 proteins (UniProt
  accessions).

Analysis plan: (1) All-protein coverage UpSet across three studies; (2) UpSet
for quiescence-regulated proteins (this study vs Paulo 2015); (3) membership
tables → `revision3/output/proteome_coverage.csv` (Table S25) and
`revision3/output/proteome_regulation_overlap.csv` (Table S26).

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
library(org.Sc.sgd.db)
library(AnnotationDbi)
library(UpSetR)

repo_root <- "/Users/david/Projects/Proteome_Rim15_Quiescence"
out_dir   <- file.path(repo_root, "revision3/output")
pub_dir   <- file.path(repo_root, "data/Publications for comparative analysis")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

is_orf <- function(x) grepl("^(Y[A-P][LR]\\d{3}[WC](-[A-Z])?|Q\\d{4})$", x)

# Map UniProt accessions → ORF (strips isoform suffix)
uniprot_to_orf <- function(accessions) {
  acc <- sub("-\\d+$", "", trimws(accessions))
  todo <- unique(acc[!is.na(acc) & acc != ""])
  m <- tryCatch(
    AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = "UNIPROT",
                          columns = "ORF") %>%
      filter(!is.na(ORF)) %>% distinct(UNIPROT, ORF) %>%
      group_by(UNIPROT) %>% dplyr::slice(1) %>% ungroup(),
    error = function(e) tibble(UNIPROT = character(), ORF = character()))
  setNames(m$ORF, m$UNIPROT)[acc]
}

# Map gene symbols → ORF
symbol_to_orf <- function(symbols) {
  key <- toupper(trimws(symbols))
  orf <- ifelse(is_orf(key), key, NA_character_)
  for (kt in c("GENENAME", "COMMON", "ALIAS")) {
    if (!kt %in% keytypes(org.Sc.sgd.db)) next
    todo <- unique(key[is.na(orf) & !is.na(key) & key != ""])
    if (length(todo) == 0) break
    m <- tryCatch(
      AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = kt,
                            columns = "ORF") %>%
        filter(!is.na(ORF)) %>% distinct(across(1), ORF) %>%
        group_by(across(1)) %>% dplyr::slice(1) %>% ungroup(),
      error = function(e) NULL)
    if (is.null(m)) next
    lookup <- setNames(m$ORF, m[[1]])
    orf[is.na(orf)] <- lookup[key[is.na(orf)]]
  }
  orf
}
```

# This study

```{r ours}
# All detected proteins — Table S5 source
s5 <- read_csv(
  file.path(repo_root, "tables/supplemental/S5_normalized_protein_ratios.csv"),
  show_col_types = FALSE)

# Leading protein accession is a single UniProt ID per row in S5
ours_detected <- s5 %>%
  transmute(accession = `Protein IDs`,
            gene      = `Gene names`,
            orf       = uniprot_to_orf(accession)) %>%
  filter(!is.na(orf)) %>%
  distinct(orf, gene)

cat("This study — detected proteins:", nrow(ours_detected),
    "| mapped to ORF:", sum(!is.na(ours_detected$orf)), "\n")

# rim15Δ-regulated proteins (q ≤ 0.05 in any contrast) — Table S16 source
de <- read_csv(
  file.path(repo_root, "tables/DE_proteins_byGenotype.csv"),
  show_col_types = FALSE)

ours_regulated <- de %>%
  filter(!is.na(qvalue), qvalue <= 0.05) %>%
  distinct(accession = `Protein IDs`, gene = `Gene names`) %>%
  mutate(orf = uniprot_to_orf(accession)) %>%
  filter(!is.na(orf))

cat("This study — rim15Δ-regulated (q≤0.05):", nrow(ours_regulated), "\n")
```

# Paulo et al. 2015

```{r paulo}
# ── All detected proteins ──────────────────────────────────────────────────
# si_001 Sheet1: header is row 9, data rows start at row 10.
# Sections separated by comment rows (first char = '#').
# Column 1 = ORF locus; ORF-like rows contain the protein data.
si1_raw <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-2_si_001.xlsx"),
  sheet    = "Sheet1",
  col_names = FALSE,
  skip     = 9)    # row 10 is the first ORF row
names(si1_raw)[1:5] <- c("locus", "fold_change", "pvalue", "gpi", "signal")

paulo_detected_orfs <- si1_raw %>%
  filter(grepl("^Y|^Q", locus, ignore.case = FALSE)) %>%
  pull(locus) %>%
  unique()
cat("Paulo 2015 — all detected ORFs:", length(paulo_detected_orfs), "\n")

paulo_detected <- tibble(orf = paulo_detected_orfs) %>%
  filter(is_orf(orf))

# ── Quiescent-enriched proteins (≥2-fold, significant) ────────────────────
# si_003 "2X or more Up in Quiescent": Accession column = ORF
si3_q <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-3_si_003.xlsx"),
  sheet = "2X or more Up in Quiescent")

paulo_quiescent <- si3_q %>%
  transmute(orf = Accession) %>%
  filter(is_orf(orf)) %>%
  distinct()
cat("Paulo 2015 — quiescent-enriched (≥2-fold):", nrow(paulo_quiescent), "\n")
```

# den Ridder et al. 2023

```{r denridder}
# mmc7 "All proteins": Accession = UniProt ID. Not a quiescence study (aerobic
# vs anaerobic in CEN.PK and IMX372 strains); included for proteome coverage
# comparison only.
dr_raw <- read_excel(
  file.path(pub_dir, "Proteome/den Ridder et al., 2023/mmc7.xlsx"),
  sheet = "All proteins")

denridder_detected <- dr_raw %>%
  transmute(accession = Accession,
            orf       = uniprot_to_orf(accession)) %>%
  filter(!is.na(orf)) %>%
  distinct(orf)
cat("den Ridder 2023 — detected proteins (UniProt→ORF):", nrow(denridder_detected),
    "| from", nrow(dr_raw), "entries\n")
```

# Coverage overlap (all detected proteins)

```{r coverage}
# ORF-to-gene lookup across all studies (prefer our own annotation)
orf_gene <- bind_rows(
  ours_detected    %>% transmute(orf, gene),
  paulo_detected   %>% mutate(gene = NA_character_),
  denridder_detected %>% mutate(gene = NA_character_)) %>%
  filter(!is.na(gene)) %>%
  group_by(orf) %>% dplyr::slice(1) %>% ungroup()

coverage <- bind_rows(
  ours_detected    %>% transmute(orf, dataset = "ThisStudy"),
  paulo_detected   %>% transmute(orf, dataset = "Paulo2015"),
  denridder_detected %>% transmute(orf, dataset = "denRidder2023")) %>%
  distinct(orf, dataset) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  # ensure all three columns exist even if a set is empty
  { if (!"ThisStudy"    %in% names(.)) mutate(., ThisStudy    = 0L) else . } %>%
  { if (!"Paulo2015"    %in% names(.)) mutate(., Paulo2015    = 0L) else . } %>%
  { if (!"denRidder2023" %in% names(.)) mutate(., denRidder2023 = 0L) else . } %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  arrange(desc(ThisStudy), desc(Paulo2015), desc(denRidder2023), orf)

write_csv(coverage, file.path(out_dir, "proteome_coverage.csv"))
cat("Coverage table written:", nrow(coverage), "unique ORFs\n")

set_order_coverage <- c("ThisStudy", "Paulo2015", "denRidder2023")
pdf(file.path(out_dir, "proteome_coverage_upset.pdf"), width = 7, height = 5)
print(upset(as.data.frame(coverage), sets = set_order_coverage,
            keep.order = TRUE, order.by = "freq",
            mainbar.y.label = "Proteins (ORF)",
            sets.x.label   = "Proteins per dataset"))
dev.off()

# Print to notebook
upset(as.data.frame(coverage), sets = set_order_coverage,
      keep.order = TRUE, order.by = "freq",
      mainbar.y.label = "Proteins (ORF)")
```

# Quiescence-regulation overlap

```{r regulation}
# Compare rim15Δ-regulated proteins (this study) with quiescent-enriched
# proteins from Paulo 2015. den Ridder is excluded here (no quiescence
# condition in that study).
regulation <- bind_rows(
  ours_regulated   %>% transmute(orf, gene, dataset = "ThisStudy_rim15reg"),
  paulo_quiescent  %>% transmute(orf, gene = NA_character_, dataset = "Paulo2015_qsnt")) %>%
  distinct(orf, dataset) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  { if (!"ThisStudy_rim15reg" %in% names(.)) mutate(., ThisStudy_rim15reg = 0L) else . } %>%
  { if (!"Paulo2015_qsnt"    %in% names(.)) mutate(., Paulo2015_qsnt    = 0L) else . } %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  mutate(category = case_when(
    ThisStudy_rim15reg == 1 & Paulo2015_qsnt == 1 ~ "shared",
    ThisStudy_rim15reg == 1                        ~ "novel_this_study",
    TRUE                                           ~ "paulo_only")) %>%
  arrange(desc(ThisStudy_rim15reg), desc(Paulo2015_qsnt), orf)

write_csv(regulation, file.path(out_dir, "proteome_regulation_overlap.csv"))
cat("Regulation overlap table written:", nrow(regulation), "unique ORFs\n")

set_order_reg <- c("ThisStudy_rim15reg", "Paulo2015_qsnt")
pdf(file.path(out_dir, "proteome_regulation_upset.pdf"), width = 6, height = 5)
print(upset(as.data.frame(regulation), sets = set_order_reg,
            keep.order = TRUE, order.by = "freq",
            mainbar.y.label = "Proteins (ORF)",
            sets.x.label   = "Proteins per set"))
dev.off()

upset(as.data.frame(regulation), sets = set_order_reg,
      keep.order = TRUE, order.by = "freq",
      mainbar.y.label = "Proteins (ORF)")
```

# Headline numbers

```{r headline}
cat("=== Coverage ===\n")
cat("This study detected:", sum(coverage$ThisStudy), "\n")
cat("Paulo 2015 detected:", sum(coverage$Paulo2015), "\n")
cat("den Ridder 2023 detected:", sum(coverage$denRidder2023), "\n")
cat("All three studies:", sum(coverage$ThisStudy & coverage$Paulo2015 & coverage$denRidder2023), "\n")
cat("ThisStudy ∩ Paulo2015:", sum(coverage$ThisStudy & coverage$Paulo2015), "\n")
cat("ThisStudy ∩ denRidder2023:", sum(coverage$ThisStudy & coverage$denRidder2023), "\n")

cat("\n=== Quiescence-regulation overlap ===\n")
cat("ThisStudy rim15Δ-regulated:", sum(regulation$ThisStudy_rim15reg), "\n")
cat("Paulo2015 quiescent-enriched:", sum(regulation$Paulo2015_qsnt), "\n")
cat("Shared:", sum(regulation$category == "shared"), "\n")
cat("Fraction of our regulated proteins in Paulo quiescent set:",
    round(sum(regulation$category == "shared") / sum(regulation$ThisStudy_rim15reg), 3), "\n")

# Top shared proteins — rename space-containing columns via base R before piping
# (S4Vectors::rename, loaded by AnnotationDbi, does not support backtick-quoted names)
cat("\nTop shared proteins (both rim15Δ-regulated and Paulo quiescent-enriched):\n")
de_slim <- de %>%
  distinct(`Protein IDs`, `Gene names`, log2FC, qvalue) %>%
  group_by(`Protein IDs`) %>%
  slice_min(qvalue, n = 1) %>%
  ungroup()
names(de_slim)[names(de_slim) == "Protein IDs"] <- "accession"
names(de_slim)[names(de_slim) == "Gene names"]  <- "gene_de"

regulation %>%
  filter(category == "shared") %>%
  left_join(de_slim, by = c("orf" = "accession")) %>%
  arrange(qvalue) %>%
  dplyr::select(orf, gene, log2FC, qvalue) %>%
  head(20) %>%
  print(n = 20)
```

```{r session}
sessionInfo()
```
