---
title: "Reconciling the mitochondrial proteome and imaging phenotypes (R2.7)"
output: html_document
---

Reviewer 2 asked us to reconcile: rim15Δ shows *increased* relative abundance of
mitochondrial proteins under carbon starvation, yet *fails to expand* mitochondrial
volume. The reconciliation argument is threefold:

1. SILAC ratios are relative (rim15Δ / WT), not absolute — an increased ratio means
   the protein is more abundant in rim15Δ *relative to WT at the same timepoint*,
   not that total mitochondrial mass increases.
2. Our coverage is partial and abundance-biased — we detect the most abundant mito
   proteins; the biogenesis/import machinery is systematically under-represented.
3. The elevated ratios are concentrated in a specific functional subset (metabolic
   enzymes: TCA, OXPHOS), not the membrane-biogenesis program (TIM/TOM, MICOS)
   that would need to expand to increase mitochondrial volume.

Reference: Morgenstern et al. 2017 (Cell Rep 19:2836), high-confidence mitochondrial
proteome with subcompartment annotation. Files in `data/mitochondrial proteome/`.

Outputs:
- `revision3/output/mito_coverage_table.csv` — n detected / n reference, per subcompartment
- `revision3/output/mito_class_log2fc.csv` — per functional class log2(rim15Δ/WT) summary
- `revision3/output/mito_coverage.pdf` — coverage bar chart
- `revision3/output/mito_by_subcompartment.pdf` — WT vs rim15Δ trajectories by sublocalization
- `revision3/output/mito_by_class.pdf` — WT vs rim15Δ trajectories by functional class
- `revision3/output/mito_class_heatmap.pdf` — per-protein log2FC heatmap by class

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(cowplot)
library(org.Sc.sgd.db)

dir.create("./revision3/output", showWarnings = FALSE, recursive = TRUE)

out_dir  <- "./revision3/output/"
mito_dir <- "./data/mitochondrial proteome/"
```

# Reference set: Morgenstern 2017 high-confidence mitochondrial proteome

```{r mito-reference}
# mmc4.xlsx sheet 2: 901 high-confidence mito proteins
# Columns: Gene Systematic Name, UniProt ID, Gene Standard Name,
#          Molecular Weight, Copy numbers (Glucose/Galactose/Glycerol),
#          Submitochondrial Localization (IMS/IM, Matrix/IM, OM, NA)
mito_ref_raw <- read_excel(
  file.path(mito_dir, "mmc4.xlsx"),
  sheet = "High-confidence mito. proteome"
)

mito_ref <- mito_ref_raw %>%
  transmute(
    orf         = `Gene Systematic Name`,
    uniprot     = `UniProt ID`,
    gene        = `Gene Standard Name1`,
    description = `Gene Description`,
    subloc_raw  = `Submitochondrial Localization2`,
    # Tidy sublocalization into three clean categories + unknown
    subloc = case_when(
      grepl("Matrix", subloc_raw, ignore.case = TRUE) ~ "Matrix/IM",
      grepl("IMS",    subloc_raw, ignore.case = TRUE) ~ "IMS/IM",
      grepl("OM",     subloc_raw, ignore.case = TRUE) ~ "OM",
      TRUE ~ "Unassigned"
    )
  )

cat("Morgenstern reference: n =", nrow(mito_ref), "proteins\n")
cat("Sublocalization breakdown:\n")
print(table(mito_ref$subloc))
```

# Functional class assignment

The reference does not provide pre-assigned functional classes. We assign them
using GO biological-process and GO cellular-component terms from org.Sc.sgd.db,
mapping to the classes most relevant to the reconciliation argument. Priority order
handles proteins that belong to multiple classes — biogenesis machinery takes
priority over metabolic function, since the reconciliation hinges on showing that
biogenesis proteins are *not* elevated.

Classes (in priority order):
1. **TIM/TOM import** — mitochondrial protein import machinery
2. **MICOS / biogenesis** — membrane contact sites and biogenesis
3. **Mitoribosome** — mitochondrial ribosome
4. **OXPHOS** — respiratory chain complexes I–V + ATP synthase
5. **TCA** — tricarboxylic acid cycle
6. **Carriers / SLC25** — mitochondrial carrier family (metabolite transport)
7. **Other mito** — all remaining reference proteins

```{r functional-class}
# GO term → class mapping (parent GO terms; children are matched via annotation)
go_classes <- tribble(
  ~go_id,        ~class,
  "GO:0005744",  "TIM/TOM",      # TIM23 translocon complex
  "GO:0005742",  "TIM/TOM",      # TIM22 complex
  "GO:0005740",  "TIM/TOM",      # outer mito membrane translocase (TOM)
  "GO:0042721",  "TIM/TOM",      # TIM54 complex
  "GO:0016021",  "TIM/TOM",      # integral membrane component (catches some TOM subunits)
  "GO:0000002",  "MICOS",        # mitochondrial genome maintenance
  "GO:0061617",  "MICOS",        # MICOS complex
  "GO:0005761",  "Mitoribosome", # mitochondrial ribosome
  "GO:0005762",  "Mitoribosome", # mitochondrial large ribosomal subunit
  "GO:0005763",  "Mitoribosome", # mitochondrial small ribosomal subunit
  "GO:0045283",  "OXPHOS",       # complex I
  "GO:0045275",  "OXPHOS",       # complex II
  "GO:0045273",  "OXPHOS",       # complex III
  "GO:0045277",  "OXPHOS",       # complex IV
  "GO:0045261",  "OXPHOS",       # complex V (ATP synthase)
  "GO:0006099",  "TCA",          # TCA cycle
  "GO:0006100",  "TCA",          # 2-oxoglutarate metabolic process
  "GO:0006839",  "Carriers",     # mitochondrial transport
  "GO:0015931",  "Carriers",     # nucleobase-containing small molecule transmembrane transport
  "GO:0008272"   , "Carriers"    # sulfate transport
)
# NOTE: we deliberately omit GO:0005741/GO:0005743 (outer/inner membrane location
# terms) from the carrier mapping — these are too broad and would misclassify
# TOM40, TIM23, and other non-carrier IM/OM proteins.

# SLC25 carrier family: keyword search on description (more specific than GO location)
# "translocator" is excluded because it also matches the TIM/TOM translocon proteins.
carrier_keywords <- c("\\bcarrier\\b", "ADP/ATP", "phosphate carrier",
                      "oxoglutarate", "dicarboxylate", "citrate carrier",
                      "carnitine", "SLC25", "ATP-Mg/Pi")

# Fetch GO-BP + GO-CC annotations for our reference ORFs
ref_orfs <- mito_ref$orf[!is.na(mito_ref$orf)]
go_ann <- AnnotationDbi::select(
  org.Sc.sgd.db,
  keys   = ref_orfs,
  keytype = "ORF",
  columns = c("GO", "ONTOLOGY")
) %>%
  filter(!is.na(GO), ONTOLOGY %in% c("BP", "CC")) %>%
  distinct(ORF, GO)

# Map GO terms to classes
go_to_class <- go_ann %>%
  inner_join(go_classes, by = c("GO" = "go_id")) %>%
  group_by(ORF) %>%
  summarise(class_go = dplyr::first(class), .groups = "drop")  # first match = highest priority

# Keyword-based carrier assignment
carrier_orfs <- mito_ref %>%
  filter(grepl(paste(carrier_keywords, collapse = "|"), description,
               ignore.case = TRUE)) %>%
  pull(orf)

# Assemble final functional class: GO-based, then keyword carrier, then Other
mito_ref <- mito_ref %>%
  left_join(go_to_class, by = c("orf" = "ORF")) %>%
  mutate(
    func_class = case_when(
      class_go %in% c("TIM/TOM","MICOS","Mitoribosome","OXPHOS","TCA") ~ class_go,
      orf %in% carrier_orfs ~ "Carriers",
      !is.na(class_go) ~ class_go,
      TRUE ~ "Other mito"
    )
  )

cat("Functional class breakdown:\n")
print(table(mito_ref$func_class))
```

# Join to detected proteome

```{r join-detected}
prot <- read_csv("./tables/master_tidy_protein.csv", show_col_types = FALSE)

# Column is "Protein.IDs" when read with read_csv (dots from spaces)
# but check the actual name first
cat("prot columns:", paste(names(prot)[1:5], collapse = ", "), "\n")
id_col <- grep("Protein", names(prot), value = TRUE)[1]
prot <- prot %>%
  mutate(leading_acc = sub(";.*$", "", .data[[id_col]]))

# Join by UniProt accession (mito_ref$uniprot ↔ prot$leading_acc)
prot_mito <- prot %>%
  inner_join(mito_ref %>% dplyr::select(orf, uniprot, gene, subloc, func_class),
             by = c("leading_acc" = "uniprot"))

cat("Reference proteins:         ", nrow(mito_ref), "\n")
cat("Detected in our study:      ", n_distinct(prot_mito$leading_acc), "\n")
cat("Overall coverage:            ",
    round(n_distinct(prot_mito$leading_acc) / nrow(mito_ref) * 100, 1), "%\n\n")

# Per-subcompartment coverage
subloc_coverage <- mito_ref %>%
  group_by(subloc) %>%
  summarise(n_reference = n(), .groups = "drop") %>%
  left_join(
    prot_mito %>% distinct(leading_acc, subloc) %>%
      count(subloc, name = "n_detected"),
    by = "subloc"
  ) %>%
  mutate(n_detected = replace_na(n_detected, 0L),
         pct_detected = round(n_detected / n_reference * 100, 1))

cat("Coverage by subcompartment:\n")
print(subloc_coverage)

write_csv(subloc_coverage, file.path(out_dir, "mito_coverage_table.csv"))
```

# Coverage plot

```{r coverage-plot}
p_coverage <- subloc_coverage %>%
  mutate(subloc = factor(subloc, levels = c("Matrix/IM","IMS/IM","OM","Unassigned"))) %>%
  ggplot(aes(x = subloc, y = pct_detected, fill = subloc)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = paste0(n_detected, " / ", n_reference)),
            vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = c("Matrix/IM" = "#4C96D7", "IMS/IM" = "#82C341",
                                "OM" = "#F5A623", "Unassigned" = "#9B9B9B")) +
  scale_y_continuous(limits = c(0, 100), expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Subcompartment", y = "Detected (%)",
       title = "Coverage of Morgenstern 2017 reference proteome") +
  theme_cowplot()

ggsave(file.path(out_dir, "mito_coverage.pdf"), p_coverage, width = 6, height = 4)
print(p_coverage)
```

# WT vs rim15Δ trajectories by subcompartment

For each nutrient (C, P) × sublocalization: mean normalized ratio (rim15Δ / WT) over
time, summarised across detected mito proteins in that subcompartment. Error bars are
±SEM across proteins. This shows whether the elevated ratios are uniform or
subcompartment-specific.

```{r subcompartment-trajectories}
# Compute per-protein mean normalized_ratio across replicates, then
# summarise per subcompartment × nutrient × time
subloc_traj <- prot_mito %>%
  filter(!is.na(normalized_ratio)) %>%
  group_by(leading_acc, gene, subloc, func_class, nutrient, samplingTime) %>%
  summarise(mean_ratio = mean(normalized_ratio, na.rm = TRUE), .groups = "drop") %>%
  group_by(subloc, nutrient, samplingTime) %>%
  summarise(
    mean_log2fc = mean(mean_ratio, na.rm = TRUE),
    sem_log2fc  = sd(mean_ratio, na.rm = TRUE) / sqrt(sum(!is.na(mean_ratio))),
    n_proteins  = sum(!is.na(mean_ratio)),
    .groups = "drop"
  ) %>%
  mutate(time_h = as.integer(sub("T", "", samplingTime)),
         subloc = factor(subloc, levels = c("Matrix/IM","IMS/IM","OM","Unassigned")),
         nutrient_label = ifelse(nutrient == "C", "Carbon starvation",
                                  "Phosphorus starvation"))

p_subloc <- ggplot(subloc_traj,
                   aes(x = time_h, y = mean_log2fc, color = subloc, group = subloc)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey60") +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = mean_log2fc - sem_log2fc,
                    ymax = mean_log2fc + sem_log2fc),
                width = 1, linewidth = 0.6, alpha = 0.8) +
  facet_wrap(~ nutrient_label, ncol = 2) +
  scale_color_manual(
    values = c("Matrix/IM" = "#4C96D7", "IMS/IM" = "#82C341",
               "OM" = "#F5A623", "Unassigned" = "#9B9B9B"),
    name = "Subcompartment"
  ) +
  scale_x_continuous(breaks = c(0, 6, 16, 30)) +
  labs(x = "Time (h)", y = expression(log[2](rim15*Delta/WT)~"(mean ± SEM across proteins)"),
       title = "Mitochondrial proteins: rim15Δ/WT by subcompartment") +
  theme_cowplot() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        legend.position = "right")

ggsave(file.path(out_dir, "mito_by_subcompartment.pdf"), p_subloc, width = 10, height = 5)
print(p_subloc)
```

# WT vs rim15Δ trajectories by functional class

The class-level trajectories provide context for the reconciliation argument.
Note: if the elevation is broadly distributed across all classes (including
biogenesis machinery), the strongest arguments remain #1 (SILAC ratios are
relative) and #2 (coverage is partial and biased toward abundant metabolic
proteins), with argument #3 stated more cautiously as "the greatest increases
are in metabolic enzymes." The figures and T30 table below show which argument
the data best supports.

```{r class-trajectories}
class_order <- c("OXPHOS","TCA","Carriers","Mitoribosome","TIM/TOM","MICOS","Other mito")
class_colors <- c("OXPHOS"       = "#D73027",
                  "TCA"          = "#FC8D59",
                  "Carriers"     = "#FEE090",
                  "Mitoribosome" = "#74ADD1",
                  "TIM/TOM"      = "#4575B4",
                  "MICOS"        = "#313695",
                  "Other mito"   = "#BDBDBD")

class_traj <- prot_mito %>%
  filter(!is.na(normalized_ratio)) %>%
  group_by(leading_acc, gene, func_class, nutrient, samplingTime) %>%
  summarise(mean_ratio = mean(normalized_ratio, na.rm = TRUE), .groups = "drop") %>%
  group_by(func_class, nutrient, samplingTime) %>%
  summarise(
    mean_log2fc = mean(mean_ratio, na.rm = TRUE),
    sem_log2fc  = sd(mean_ratio, na.rm = TRUE) / sqrt(sum(!is.na(mean_ratio))),
    n_proteins  = sum(!is.na(mean_ratio)),
    .groups = "drop"
  ) %>%
  mutate(time_h = as.integer(sub("T", "", samplingTime)),
         func_class = factor(func_class, levels = class_order),
         nutrient_label = ifelse(nutrient == "C", "Carbon starvation",
                                  "Phosphorus starvation"))

p_class <- ggplot(class_traj,
                  aes(x = time_h, y = mean_log2fc, color = func_class, group = func_class)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey60") +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = mean_log2fc - sem_log2fc,
                    ymax = mean_log2fc + sem_log2fc),
                width = 1, linewidth = 0.6, alpha = 0.8) +
  facet_wrap(~ nutrient_label, ncol = 2) +
  scale_color_manual(values = class_colors, name = "Functional class") +
  scale_x_continuous(breaks = c(0, 6, 16, 30)) +
  labs(x = "Time (h)", y = expression(log[2](rim15*Delta/WT)~"(mean ± SEM across proteins)"),
       title = "Mitochondrial proteins: rim15Δ/WT by functional class") +
  theme_cowplot() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        legend.position = "right")

ggsave(file.path(out_dir, "mito_by_class.pdf"), p_class, width = 10, height = 5)
print(p_class)
```

# Per-class summary table (response letter numbers)

```{r class-table}
class_log2fc <- prot_mito %>%
  filter(!is.na(normalized_ratio), nutrient == "C") %>%
  group_by(leading_acc, orf, gene, func_class, subloc, samplingTime) %>%
  summarise(mean_ratio = mean(normalized_ratio, na.rm = TRUE), .groups = "drop") %>%
  group_by(func_class, subloc, samplingTime) %>%
  summarise(
    mean_log2fc  = round(mean(mean_ratio, na.rm = TRUE), 3),
    sem_log2fc   = round(sd(mean_ratio, na.rm = TRUE) / sqrt(sum(!is.na(mean_ratio))), 3),
    n_proteins   = sum(!is.na(mean_ratio)),
    .groups = "drop"
  ) %>%
  mutate(time_h = as.integer(sub("T", "", samplingTime))) %>%
  arrange(func_class, time_h)

write_csv(class_log2fc, file.path(out_dir, "mito_class_log2fc.csv"))

# Print T30 (peak starvation) carbon for the response letter
cat("=== Carbon starvation T30: mean log2(rim15Δ/WT) per class ===\n")
class_log2fc %>%
  filter(time_h == 30) %>%
  dplyr::select(func_class, subloc, mean_log2fc, sem_log2fc, n_proteins) %>%
  arrange(desc(mean_log2fc)) %>%
  print()
```

# Heatmap: per-protein log2FC at T30, carbon, ordered by class

```{r class-heatmap, fig.height=10, fig.width=7}
heatmap_dat <- prot_mito %>%
  filter(!is.na(normalized_ratio), nutrient == "C", samplingTime == "T30") %>%
  group_by(leading_acc, gene, func_class, subloc) %>%
  summarise(mean_ratio = mean(normalized_ratio, na.rm = TRUE), .groups = "drop") %>%
  mutate(gene_label = coalesce(gene, leading_acc),
         func_class = factor(func_class, levels = class_order)) %>%
  arrange(func_class, desc(mean_ratio))

heatmap_dat$gene_label <- factor(heatmap_dat$gene_label,
                                  levels = heatmap_dat$gene_label)

p_heat <- ggplot(heatmap_dat,
                 aes(x = 1, y = gene_label, fill = mean_ratio)) +
  geom_tile() +
  facet_grid(func_class ~ ., scales = "free_y", space = "free_y",
             switch = "y") +
  scale_fill_gradient2(low = "#4575B4", mid = "white", high = "#D73027",
                       midpoint = 0, limits = c(-2, 2), oob = scales::squish,
                       name = expression(log[2](rim15*Delta/WT))) +
  labs(title = "Mitochondrial proteins: carbon starvation T30 h",
       y = NULL, x = NULL) +
  theme_cowplot(font_size = 8) +
  theme(axis.text.x    = element_blank(),
        axis.ticks.x   = element_blank(),
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0, hjust = 1, size = 8),
        strip.background  = element_blank(),
        legend.position   = "right")

ggsave(file.path(out_dir, "mito_class_heatmap.pdf"), p_heat,
       width = 7, height = max(4, nrow(heatmap_dat) * 0.12 + 1))
print(p_heat)
```

# Sanity checks

```{r sanity}
cat("=== Overall ===\n")
cat("Morgenstern reference (high-confidence):", nrow(mito_ref), "\n")
cat("Detected in our study:", n_distinct(prot_mito$leading_acc), "/", nrow(mito_ref),
    sprintf("(%.1f%%)\n", n_distinct(prot_mito$leading_acc) / nrow(mito_ref) * 100))

# Well-known mito markers should be detected
markers <- c("COX1","ATP2","SDH1","IDH1","TIM23","TOM40","MRPL32","CTP1")
cat("\nExpected markers (should all be in prot_mito):\n")
print(mito_ref %>%
  filter(gene %in% markers) %>%
  left_join(prot_mito %>% distinct(leading_acc, nutrient, samplingTime),
            by = c("uniprot" = "leading_acc")) %>%
  distinct(gene, subloc, func_class, nutrient) %>%
  arrange(gene))
```

```{r session}
sessionInfo()
```
