---
title: "Reconciling the mitochondrial proteome and imaging phenotypes (R2.7)"
output: html_document
---

Reviewer 2 asked us to reconcile two statements: rim15KO shows *increased*
relative abundance of mitochondrial proteins under carbon starvation, yet
*fails to expand* mitochondrial volume. The reconciliation to demonstrate:
SILAC ratios are relative not absolute; mitochondrial coverage is partial and
abundance-biased; and the increase is confined to a specific functional subset
(carriers / metabolic enzymes), not the membrane-biogenesis program.

Deliverables: coverage numbers (n detected / n annotated, overall and per
subcompartment), WT vs rim15KO plots by subcompartment/functional class, and a
supplementary table of per-class log2(rim15KO/WT).

```{r setup, message=FALSE}
library(tidyverse)

dir.create("./revision3/output", showWarnings = FALSE, recursive = TRUE)

# Detected proteome (tidy long format, both genotypes)
prot <- read_csv("./tables/master_tidy_protein.csv")
```

```{r mito-reference}
# Reference mitochondrial proteome with subcompartment annotation:
# Morgenstern et al. 2017 (Cell Rep 19:2836) high-confidence mito proteome,
# Supplementary Table S1 - includes subcompartment (matrix, IM, OM, IMS)
# and functional class. Download once into data/published/:
#   https://www.cell.com/cms/10.1016/j.celrep.2017.06.014/attachment/…/mmc2.xlsx
# Fallback: GO:0005739 annotation from org.Sc.sgd.db.
#
# TODO: load reference, join to detected proteins by ORF.
# Classes to assign per the response plan: matrix, inner membrane (OXPHOS
# I-V), outer membrane, IMS, mitoribosome, TCA, carriers/SLC25, TIM/TOM
# import, MICOS/biogenesis.
```

```{r coverage-and-classes}
# 1. Coverage: n detected / n reference, overall + per subcompartment
#    -> revision3/output/mito_coverage_table.csv  (feeds "[N] of [M]" in the
#       response letter and Results).
# 2. Per-class WT vs rim15KO trajectories (carbon and phosphorus):
#    -> revision3/output/mito_by_class.pdf
# 3. Per-class log2(rim15KO/WT) summary table
#    -> revision3/output/mito_class_log2fc.csv
```

```{r session}
sessionInfo()
```
