---
title: "Cross-dataset comparison of quiescence proteome — Paulo et al. 2015 (R2.3)"
output: html_document
---

Reviewer 2 requested a comparison of our detected proteins and starvation-regulated
proteins with those in published quiescence proteome studies. This script compares
our carbon-starvation time course (WT cells, Table S5) with Paulo et al. 2015
(Mol Cell Proteomics 14:1998–2015), which used sedimentation fractionation to
compare quiescent vs log-phase *S. cerevisiae* proteomes.

**Source tables:**

- **This study** — `tables/supplemental/S5_normalized_protein_ratios.csv` (Table S5).
  All detected proteins = any protein with a quantified WT ratio in carbon starvation.
  Regulated proteins = WT log₂FC at T30 vs T00 in carbon starvation, threshold
  |log₂FC| > 1 (≥2-fold), matching Paulo et al. 2015.
- **Paulo et al. 2015** — `NIHMS464126-supplement-2_si_001.xlsx` Sheet1 (all 2438
  detected ORFs) and `NIHMS464126-supplement-3_si_003.xlsx` (quiescent-enriched
  and log-enriched protein lists, both ≥2-fold).

Outputs → `revision3/output/proteome_coverage.csv` (Table S25) and
`revision3/output/proteome_regulation_overlap.csv` (Table S26).

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
library(org.Sc.sgd.db)
library(AnnotationDbi)
library(UpSetR)

repo_root <- "/Users/david/Projects/Proteome_Rim15_Quiescence"
out_dir   <- file.path(repo_root, "revision3/output")
pub_dir   <- file.path(repo_root, "data/Publications for comparative analysis")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

is_orf <- function(x) grepl("^(Y[A-P][LR]\\d{3}[WC](-[A-Z])?|Q\\d{4})$", x)

uniprot_to_orf <- function(accessions) {
  acc <- sub("-\\d+$", "", trimws(accessions))
  todo <- unique(acc[!is.na(acc) & acc != ""])
  m <- tryCatch(
    AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = "UNIPROT",
                          columns = "ORF") %>%
      filter(!is.na(ORF)) %>% distinct(UNIPROT, ORF) %>%
      group_by(UNIPROT) %>% dplyr::slice(1) %>% ungroup(),
    error = function(e) tibble(UNIPROT = character(), ORF = character()))
  setNames(m$ORF, m$UNIPROT)[acc]
}
```

# This study — carbon starvation, WT

```{r ours}
# Table S5 source: tables/supplemental/S5_normalized_protein_ratios.csv
# Base-R rename before piping (avoids S4Vectors::rename mask from AnnotationDbi)
s5 <- read_csv(
  file.path(repo_root, "tables/supplemental/S5_normalized_protein_ratios.csv"),
  show_col_types = FALSE)
names(s5)[names(s5) == "Protein IDs"] <- "accession"
names(s5)[names(s5) == "Gene names"]  <- "gene"

# All detected: any protein with ≥1 non-NA WT carbon starvation ratio
wt_c_cols <- grep("^WT_C_T", names(s5), value = TRUE)
ours_detected <- s5 %>%
  filter(rowSums(!is.na(across(all_of(wt_c_cols)))) >= 1) %>%
  transmute(accession, gene, orf = uniprot_to_orf(accession)) %>%
  filter(!is.na(orf)) %>%
  distinct(orf, gene)

cat("This study — proteins detected in C starvation (WT):", nrow(ours_detected), "\n")

# Starvation-regulated: WT log2FC at T30 vs T00 in C starvation
# Require ≥2 non-NA replicates at both timepoints; threshold |log2FC| > 1
t00_cols <- grep("^WT_C_T00_R", names(s5), value = TRUE)
t30_cols <- grep("^WT_C_T30_R", names(s5), value = TRUE)

fc_C <- s5 %>%
  transmute(
    accession, gene,
    mean_t00 = rowMeans(log2(across(all_of(t00_cols))), na.rm = TRUE),
    mean_t30 = rowMeans(log2(across(all_of(t30_cols))), na.rm = TRUE),
    n_t00    = rowSums(!is.na(across(all_of(t00_cols)))),
    n_t30    = rowSums(!is.na(across(all_of(t30_cols)))),
    log2FC   = mean_t30 - mean_t00
  ) %>%
  filter(n_t00 >= 2, n_t30 >= 2, is.finite(log2FC)) %>%
  mutate(orf = uniprot_to_orf(accession)) %>%
  filter(!is.na(orf))

ours_up   <- fc_C %>% filter(log2FC >  1)
ours_down <- fc_C %>% filter(log2FC < -1)

cat("C starvation — proteins with FC data:", nrow(fc_C), "\n")
cat("  upregulated (≥2-fold at T30):", nrow(ours_up), "\n")
cat("  downregulated (≥2-fold at T30):", nrow(ours_down), "\n")
```

# Paulo et al. 2015

```{r paulo}
# All detected proteins — Sheet1, all ORF-like rows across all sections
si1_raw <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-2_si_001.xlsx"),
  sheet = "Sheet1", col_names = FALSE, skip = 9)
names(si1_raw)[1] <- "locus"

paulo_detected <- si1_raw %>%
  filter(grepl("^Y|^Q", locus)) %>%
  transmute(orf = locus) %>%
  filter(is_orf(orf)) %>%
  distinct()
cat("Paulo 2015 — all detected:", nrow(paulo_detected), "\n")

# Quiescent-enriched (≥2-fold up in quiescent cells)
paulo_q_up <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-3_si_003.xlsx"),
  sheet = "2X or more Up in Quiescent") %>%
  transmute(orf = Accession) %>% filter(is_orf(orf)) %>% distinct()
cat("Paulo 2015 — quiescent-enriched:", nrow(paulo_q_up), "\n")

# Log-enriched (≥2-fold up in log phase = decreasing during quiescence)
paulo_log_up <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-3_si_003.xlsx"),
  sheet = "2X or more Up in LOG") %>%
  transmute(orf = Accession) %>% filter(is_orf(orf)) %>% distinct()
cat("Paulo 2015 — log-enriched (down in quiescence):", nrow(paulo_log_up), "\n")
```

# Coverage overlap

```{r coverage}
orf_gene <- ours_detected %>% filter(!is.na(gene)) %>%
  group_by(orf) %>% dplyr::slice(1) %>% ungroup()

coverage <- bind_rows(
  ours_detected  %>% transmute(orf, dataset = "ThisStudy"),
  paulo_detected %>% transmute(orf, dataset = "Paulo2015")) %>%
  distinct(orf, dataset) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  { if (!"ThisStudy"  %in% names(.)) mutate(., ThisStudy  = 0L) else . } %>%
  { if (!"Paulo2015"  %in% names(.)) mutate(., Paulo2015  = 0L) else . } %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  arrange(desc(ThisStudy), desc(Paulo2015), orf)

write_csv(coverage, file.path(out_dir, "proteome_coverage.csv"))

n_ours    <- sum(coverage$ThisStudy)
n_paulo   <- sum(coverage$Paulo2015)
n_overlap <- sum(coverage$ThisStudy & coverage$Paulo2015)
n_ours_only  <- sum(coverage$ThisStudy & !coverage$Paulo2015)
n_paulo_only <- sum(!coverage$ThisStudy & coverage$Paulo2015)

cat(sprintf("Coverage: ThisStudy=%d  Paulo2015=%d  overlap=%d (%.0f%% of ours)\n",
            n_ours, n_paulo, n_overlap, 100 * n_overlap / n_ours))

pdf(file.path(out_dir, "proteome_coverage_upset.pdf"), width = 6, height = 4)
print(upset(as.data.frame(coverage), sets = c("ThisStudy", "Paulo2015"),
            keep.order = TRUE, order.by = "freq",
            mainbar.y.label = "Proteins", sets.x.label = "Proteins detected"))
dev.off()
upset(as.data.frame(coverage), sets = c("ThisStudy", "Paulo2015"),
      keep.order = TRUE, order.by = "freq", mainbar.y.label = "Proteins")
```

# Starvation-regulation overlap

```{r regulation}
regulation <- bind_rows(
  ours_up      %>% transmute(orf, dataset = "ThisStudy_up"),
  ours_down    %>% transmute(orf, dataset = "ThisStudy_down"),
  paulo_q_up   %>% transmute(orf, dataset = "Paulo_q_up"),
  paulo_log_up %>% transmute(orf, dataset = "Paulo_log_up")) %>%
  distinct(orf, dataset) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  { if (!"ThisStudy_up"   %in% names(.)) mutate(., ThisStudy_up   = 0L) else . } %>%
  { if (!"ThisStudy_down" %in% names(.)) mutate(., ThisStudy_down = 0L) else . } %>%
  { if (!"Paulo_q_up"     %in% names(.)) mutate(., Paulo_q_up     = 0L) else . } %>%
  { if (!"Paulo_log_up"   %in% names(.)) mutate(., Paulo_log_up   = 0L) else . } %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  arrange(desc(ThisStudy_up), desc(ThisStudy_down),
          desc(Paulo_q_up), desc(Paulo_log_up), orf)

write_csv(regulation, file.path(out_dir, "proteome_regulation_overlap.csv"))

# ── Detailed overlap table ───────────────────────────────────────────────────
# Proteins detected in both studies, with our log2FC (T30/T00 WT C starvation)
# alongside Paulo 2015 quantitative metrics (ANOVA p-value and normalized
# spectral-count log2 ratio; "ratio change 1/2" is Q-dense/Log in Paulo 2015).
paulo_q_detail <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-3_si_003.xlsx"),
  sheet = "2X or more Up in Quiescent") %>%
  filter(is_orf(Accession)) %>%
  transmute(orf = Accession,
            paulo_pvalue    = `anova p value`,
            log2FC_Paulo    = as.numeric(`norm spec count log(2) ratio change 1/2`),
            paulo_category  = "quiescent_enriched")

paulo_l_detail <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-3_si_003.xlsx"),
  sheet = "2X or more Up in LOG") %>%
  filter(is_orf(Accession)) %>%
  transmute(orf = Accession,
            paulo_pvalue    = `anova p value`,
            log2FC_Paulo    = as.numeric(`norm spec count log(2) ratio change 1/2`),
            paulo_category  = "log_enriched")

paulo_detail <- bind_rows(paulo_q_detail, paulo_l_detail) %>%
  distinct(orf, .keep_all = TRUE)

# All proteins detected in both studies, annotated with direction + metrics
overlap_detailed <- fc_C %>%
  filter(orf %in% paulo_detected$orf) %>%
  transmute(orf, gene, log2FC_thisStudy = log2FC) %>%
  left_join(paulo_detail, by = "orf") %>%
  mutate(paulo_category = replace_na(paulo_category, "detected_only")) %>%
  arrange(paulo_category, desc(abs(log2FC_thisStudy)))

write_csv(overlap_detailed, file.path(out_dir, "proteome_overlap_detailed.csv"))
cat(sprintf("Detailed overlap: %d proteins in both studies\n", nrow(overlap_detailed)))
cat(sprintf("  quiescent_enriched: %d  |  log_enriched: %d  |  detected_only: %d\n",
            sum(overlap_detailed$paulo_category == "quiescent_enriched"),
            sum(overlap_detailed$paulo_category == "log_enriched"),
            sum(overlap_detailed$paulo_category == "detected_only")))
cat(sprintf("  Paulo log2FC (ratio change 1/2) available for: %d proteins\n",
            sum(!is.na(overlap_detailed$log2FC_Paulo))))

n_up_shared   <- sum(regulation$ThisStudy_up   & regulation$Paulo_q_up)
n_down_shared <- sum(regulation$ThisStudy_down & regulation$Paulo_log_up)

cat(sprintf("Up overlap   (ThisStudy_up ∩ Paulo_q_up):    %d / %d (%.0f%%)\n",
            n_up_shared,   sum(regulation$ThisStudy_up),
            100 * n_up_shared   / sum(regulation$ThisStudy_up)))
cat(sprintf("Down overlap (ThisStudy_down ∩ Paulo_log_up): %d / %d (%.0f%%)\n",
            n_down_shared, sum(regulation$ThisStudy_down),
            100 * n_down_shared / sum(regulation$ThisStudy_down)))

set_reg <- c("ThisStudy_up", "ThisStudy_down", "Paulo_q_up", "Paulo_log_up")
pdf(file.path(out_dir, "proteome_regulation_upset.pdf"), width = 8, height = 5)
print(upset(as.data.frame(regulation), sets = set_reg, keep.order = TRUE,
            order.by = "freq", mainbar.y.label = "Proteins",
            sets.x.label = "Proteins per set"))
dev.off()
upset(as.data.frame(regulation), sets = set_reg, keep.order = TRUE,
      order.by = "freq", mainbar.y.label = "Proteins")
```

# Statistical tests (Fisher's exact / hypergeometric)

```{r statistics}
# Universe for all tests = proteins with T30/T00 fold-change data in our
# C starvation WT data (fc_C). These are the proteins that could potentially
# be called regulated in our study; Paulo set sizes are conditioned on
# those proteins overlapping our detectable universe.

N <- nrow(fc_C)   # total testable proteins in our study

# Helper: run Fisher's exact test and return a tidy summary.
# Contingency table:
#   [ a  b ]   a = in both sets
#   [ c  d ]   b = in set2 only
#              c = in set1 only
#              d = in neither
fisher_overlap <- function(set1_orfs, set2_orfs, universe_orfs, label) {
  a <- sum(universe_orfs %in% set1_orfs & universe_orfs %in% set2_orfs)
  b <- sum(!universe_orfs %in% set1_orfs & universe_orfs %in% set2_orfs)
  c <- sum(universe_orfs %in% set1_orfs & !universe_orfs %in% set2_orfs)
  d <- sum(!universe_orfs %in% set1_orfs & !universe_orfs %in% set2_orfs)
  ct <- chisq.test(matrix(c(a, c, b, d), nrow = 2))
  tibble(comparison = label,
         n_set1 = a + c, n_set2 = a + b, n_overlap = a, N = N,
         chi_sq = ct$statistic, p_value = ct$p.value)
}

universe <- fc_C$orf   # all proteins with quantifiable FC in our C starvation data

tests <- bind_rows(
  fisher_overlap(ours_up$orf,   paulo_q_up$orf,   universe,
                 "ThisStudy_up vs Paulo_q_up (both increase in quiescence)"),
  fisher_overlap(ours_down$orf, paulo_log_up$orf, universe,
                 "ThisStudy_down vs Paulo_log_up (both decrease in quiescence)")
)

print(tests)

# Coverage test: is the detection overlap greater than expected?
# Universe = union of both studies (total unique ORFs detected)
N_cov <- nrow(coverage)
ct_cov <- chisq.test(
  matrix(c(n_overlap,
           n_paulo_only,
           n_ours_only,
           N_cov - n_overlap - n_paulo_only - n_ours_only),
         nrow = 2))
cat(sprintf("\nCoverage overlap chi-square test: X²=%.1f, p=%s\n",
            ct_cov$statistic,
            formatC(ct_cov$p.value, format = "e", digits = 2)))
```

# GO enrichment of doubly-regulated proteins

```{r go_enrichment, message=FALSE}
library(GO.db)

# Universe = proteins detected in both this study and Paulo 2015
universe_orfs <- overlap_detailed$orf

# Doubly-regulated sets
up_orfs   <- regulation %>% filter(ThisStudy_up == 1,   Paulo_q_up == 1)   %>% pull(orf)
down_orfs <- regulation %>% filter(ThisStudy_down == 1, Paulo_log_up == 1) %>% pull(orf)

cat("Up overlap (ThisStudy_up ∩ Paulo_q_up):",   length(up_orfs),   "proteins\n")
cat("Down overlap (ThisStudy_down ∩ Paulo_log_up):", length(down_orfs), "proteins\n")
cat("Universe (detected in both studies):",       length(universe_orfs), "proteins\n\n")

# GO BP annotations for the universe
go_annot <- AnnotationDbi::select(
  org.Sc.sgd.db, keys = universe_orfs, keytype = "ORF",
  columns = c("GO", "ONTOLOGY")) %>%
  filter(ONTOLOGY == "BP", !is.na(GO)) %>%
  distinct(ORF, GO)

# GO term descriptions
go_desc <- AnnotationDbi::select(
  GO.db, keys = unique(go_annot$GO),
  keytype = "GOID", columns = "TERM") %>%
  distinct(GOID, TERM)

# Hypergeometric enrichment: are genes in gene_set over-represented in each GO term?
# phyper(x-1, m, N-m, k, lower.tail=FALSE):
#   x = overlap, m = term size in universe, N = universe size, k = gene-set size
go_hyper <- function(gene_set, universe_annot, go_desc, min_size = 3, max_size = 500) {
  N <- length(unique(universe_annot$ORF))
  k <- length(gene_set)
  map_dfr(unique(universe_annot$GO), function(term) {
    in_term <- unique(universe_annot$ORF[universe_annot$GO == term])
    m <- length(in_term)
    if (m < min_size || m > max_size) return(NULL)
    x <- sum(gene_set %in% in_term)
    if (x == 0) return(NULL)
    p <- phyper(x - 1L, m, N - m, k, lower.tail = FALSE)
    tibble(GO = term, n_term = m, n_overlap = x,
           fold_enrichment = round((x / k) / (m / N), 2),
           p_value = p)
  }) %>%
    left_join(go_desc, by = c("GO" = "GOID")) %>%
    mutate(p_adj = p.adjust(p_value, method = "BH"),
           gene_set_n = k, universe_n = N) %>%
    filter(p_adj < 0.05) %>%
    arrange(p_adj) %>%
    dplyr::select(GO, TERM, n_term, n_overlap, fold_enrichment,
                  p_value, p_adj, gene_set_n, universe_n)
}

go_up   <- go_hyper(up_orfs,   go_annot, go_desc)
go_down <- go_hyper(down_orfs, go_annot, go_desc)

cat("GO BP terms significant at p_adj < 0.05:\n")
cat("  Up overlap:", nrow(go_up), "terms\n")
cat("  Down overlap:", nrow(go_down), "terms\n\n")

if (nrow(go_up)   > 0) { cat("--- Up ---\n");   print(go_up   %>% head(20)) }
if (nrow(go_down) > 0) { cat("--- Down ---\n"); print(go_down %>% head(20)) }

write_csv(go_up,   file.path(out_dir, "go_up_overlap.csv"))
write_csv(go_down, file.path(out_dir, "go_down_overlap.csv"))
```

# Summary paragraph

```{r paragraph, results='asis'}
pct_cov  <- round(100 * n_overlap / n_ours)
pct_up   <- round(100 * n_up_shared   / sum(regulation$ThisStudy_up))
pct_down <- round(100 * n_down_shared / sum(regulation$ThisStudy_down))
p_up    <- tests$p_value[1]
p_down  <- tests$p_value[2]
chi_up  <- round(tests$chi_sq[1], 1)
chi_down <- round(tests$chi_sq[2], 1)

fmt_p <- function(p) {
  if (p < 0.001) formatC(p, format = "e", digits = 1) else round(p, 3)
}

cat(sprintf(
"To evaluate the comprehensiveness of our proteome, we compared the proteins
detected in our carbon-starvation time course with those reported by Paulo et al.
(2015), who used sedimentation fractionation to characterise the quiescent
*S. cerevisiae* proteome. Of the %d proteins we detected in WT cells during
carbon starvation, %d (%d%%) were also detected by Paulo et al. (2015), who
identified %d proteins in total. The %d proteins detected only by Paulo et al.
likely reflect the greater depth of label-free quantification relative to
SILAC-based approaches. To determine whether the proteins regulated during
carbon starvation in our study are consistent with those characterised as
quiescence-specific by Paulo et al. (2015), we applied Fisher's exact test.
Of proteins that increased ≥2-fold in abundance at 30 h in our study, %d of %d
(%d%%) were also enriched in quiescent cells by Paulo et al. (2015)
(χ² = %.1f, p = %s). Similarly, %d of %d proteins that decreased ≥2-fold
(%d%%) were enriched in log-phase cells by Paulo et al. (2015) — i.e., they
decrease during quiescence entry in both studies (χ² = %.1f, p = %s).
Together, these results demonstrate that our approach detects a representative
subset of the quiescence proteome and that starvation-induced abundance changes
are directionally consistent with an independent characterisation of quiescent
yeast.",
  n_ours, n_overlap, pct_cov, n_paulo, n_paulo_only,
  n_up_shared,   sum(regulation$ThisStudy_up),   pct_up,   chi_up,   fmt_p(p_up),
  n_down_shared, sum(regulation$ThisStudy_down), pct_down, chi_down, fmt_p(p_down)))
```

```{r session}
sessionInfo()
```
