---
title: "Cross-dataset comparison of quiescence proteome (R2.3)"
output: html_document
---

Reviewer 2 requested a systematic comparison of our starvation-regulated proteins
with those reported in published quiescence proteome studies.

**Source tables from this study:**

- **Detected proteins / fold-change computation** —
  `tables/supplemental/S5_normalized_protein_ratios.csv` (Table S5): all
  proteins with quantified SILAC ratios. For each protein, the WT log₂ fold
  change at 30 h vs 0 h is computed separately for carbon (C) and nitrogen (N)
  starvation. Proteins with mean |log₂FC| > 1 across replicates (≥2-fold) are
  called regulated, matching the Paulo et al. 2015 threshold.

**Published datasets:**

- **Paulo et al. 2015** (Mol Cell Proteomics 14:1998–2015,
  `Proteome/Paulo et al., 2015/`) — label-free quant of quiescent vs log-phase
  *S. cerevisiae* (sedimentation fractionation). Three outputs used from
  `NIHMS464126-supplement-2_si_001.xlsx` (Sheet1, all detected ORFs) and
  `NIHMS464126-supplement-3_si_003.xlsx`:
  (a) all detected proteins (Sheet1, 2438 ORFs);
  (b) proteins ≥2-fold enriched in quiescent cells ("2X or more Up in Quiescent", 235 proteins);
  (c) proteins ≥2-fold enriched in log-phase cells ("2X or more Up in LOG", 332 proteins —
  i.e., decreasing during quiescence entry).
- **den Ridder et al. 2023** (Biotechnol Biofuels 16, `Proteome/den Ridder et al., 2023/`)
  — TMT proteome of *S. cerevisiae* under aerobic/anaerobic conditions (not a
  quiescence study; included for proteome coverage comparison only, `mmc7.xlsx`).

Analysis: (1) Detected protein coverage UpSet across three studies; (2) UpSet
and membership table comparing starvation-regulated proteins from this study
(C and N starvation, WT cells, T30 vs T00) with Paulo 2015 quiescent-enriched
and log-enriched sets → `revision3/output/proteome_coverage.csv` (Table S25)
and `revision3/output/proteome_regulation_overlap.csv` (Table S26).

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
library(org.Sc.sgd.db)
library(AnnotationDbi)
library(UpSetR)

repo_root <- "/Users/david/Projects/Proteome_Rim15_Quiescence"
out_dir   <- file.path(repo_root, "revision3/output")
pub_dir   <- file.path(repo_root, "data/Publications for comparative analysis")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

is_orf <- function(x) grepl("^(Y[A-P][LR]\\d{3}[WC](-[A-Z])?|Q\\d{4})$", x)

uniprot_to_orf <- function(accessions) {
  acc <- sub("-\\d+$", "", trimws(accessions))
  todo <- unique(acc[!is.na(acc) & acc != ""])
  m <- tryCatch(
    AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = "UNIPROT",
                          columns = "ORF") %>%
      filter(!is.na(ORF)) %>% distinct(UNIPROT, ORF) %>%
      group_by(UNIPROT) %>% dplyr::slice(1) %>% ungroup(),
    error = function(e) tibble(UNIPROT = character(), ORF = character()))
  setNames(m$ORF, m$UNIPROT)[acc]
}

symbol_to_orf <- function(symbols) {
  key <- toupper(trimws(symbols))
  orf <- ifelse(is_orf(key), key, NA_character_)
  for (kt in c("GENENAME", "COMMON", "ALIAS")) {
    if (!kt %in% keytypes(org.Sc.sgd.db)) next
    todo <- unique(key[is.na(orf) & !is.na(key) & key != ""])
    if (length(todo) == 0) break
    m <- tryCatch(
      AnnotationDbi::select(org.Sc.sgd.db, keys = todo, keytype = kt,
                            columns = "ORF") %>%
        filter(!is.na(ORF)) %>% distinct(across(1), ORF) %>%
        group_by(across(1)) %>% dplyr::slice(1) %>% ungroup(),
      error = function(e) NULL)
    if (is.null(m)) next
    lookup <- setNames(m$ORF, m[[1]])
    orf[is.na(orf)] <- lookup[key[is.na(orf)]]
  }
  orf
}
```

# This study — detected proteins and starvation fold changes

```{r ours}
# Table S5 source: tables/supplemental/S5_normalized_protein_ratios.csv
s5 <- read_csv(
  file.path(repo_root, "tables/supplemental/S5_normalized_protein_ratios.csv"),
  show_col_types = FALSE)

# Map to ORF — base-R rename before piping to avoid S4Vectors::rename mask
names(s5)[names(s5) == "Protein IDs"] <- "accession"
names(s5)[names(s5) == "Gene names"]  <- "gene"

ours_detected <- s5 %>%
  transmute(accession, gene,
            orf = uniprot_to_orf(accession)) %>%
  filter(!is.na(orf)) %>%
  distinct(orf, gene)

cat("This study — detected proteins:", nrow(ours_detected), "\n")

# ── Compute WT log2FC at T30 vs T00 for C and N starvation ─────────────────
# Each cell in S5 is the normalized SILAC ratio (protein / spike-in).
# log2(T30/T00) = log2(ratio_T30) - log2(ratio_T00), computed per replicate,
# then averaged. Require ≥ 2 non-NA replicates at both timepoints.
# Threshold |log2FC| > 1 (≥2-fold) matches Paulo et al. 2015.

compute_fc <- function(df, nutrient) {
  t00_cols <- grep(sprintf("^WT_%s_T00_R", nutrient), names(df), value = TRUE)
  t30_cols <- grep(sprintf("^WT_%s_T30_R", nutrient), names(df), value = TRUE)
  df %>%
    transmute(
      accession, gene,
      mean_t00 = rowMeans(log2(across(all_of(t00_cols))), na.rm = TRUE),
      mean_t30 = rowMeans(log2(across(all_of(t30_cols))), na.rm = TRUE),
      n_t00    = rowSums(!is.na(across(all_of(t00_cols)))),
      n_t30    = rowSums(!is.na(across(all_of(t30_cols)))),
      log2FC   = mean_t30 - mean_t00
    ) %>%
    filter(n_t00 >= 2, n_t30 >= 2, is.finite(log2FC)) %>%
    mutate(
      nutrient  = nutrient,
      direction = case_when(
        log2FC >  1 ~ "up",
        log2FC < -1 ~ "down",
        TRUE        ~ "unchanged"
      )
    )
}

fc_C <- compute_fc(s5, "C")
fc_N <- compute_fc(s5, "N")

cat("C starvation proteins with FC data:", nrow(fc_C),
    "| up:", sum(fc_C$direction == "up"),
    "| down:", sum(fc_C$direction == "down"), "\n")
cat("N starvation proteins with FC data:", nrow(fc_N),
    "| up:", sum(fc_N$direction == "up"),
    "| down:", sum(fc_N$direction == "down"), "\n")

# Regulated in at least one condition
ours_up   <- bind_rows(
  fc_C %>% filter(direction == "up")   %>% mutate(condition = "C"),
  fc_N %>% filter(direction == "up")   %>% mutate(condition = "N")) %>%
  mutate(orf = uniprot_to_orf(accession)) %>% filter(!is.na(orf)) %>%
  distinct(orf, .keep_all = TRUE)

ours_down <- bind_rows(
  fc_C %>% filter(direction == "down") %>% mutate(condition = "C"),
  fc_N %>% filter(direction == "down") %>% mutate(condition = "N")) %>%
  mutate(orf = uniprot_to_orf(accession)) %>% filter(!is.na(orf)) %>%
  distinct(orf, .keep_all = TRUE)

cat("\nThis study — upregulated (≥2-fold, C or N):", nrow(ours_up), "\n")
cat("This study — downregulated (≥2-fold, C or N):", nrow(ours_down), "\n")
```

# Paulo et al. 2015

```{r paulo}
# ── All detected proteins ──────────────────────────────────────────────────
si1_raw <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-2_si_001.xlsx"),
  sheet = "Sheet1", col_names = FALSE, skip = 9)
names(si1_raw)[1] <- "locus"

paulo_detected <- si1_raw %>%
  filter(grepl("^Y|^Q", locus)) %>%
  transmute(orf = locus) %>%
  filter(is_orf(orf)) %>%
  distinct()
cat("Paulo 2015 — all detected:", nrow(paulo_detected), "\n")

# ── Quiescent-enriched (≥2-fold up in quiescent cells) ────────────────────
si3_q <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-3_si_003.xlsx"),
  sheet = "2X or more Up in Quiescent")
paulo_q_up <- si3_q %>%
  transmute(orf = Accession) %>%
  filter(is_orf(orf)) %>%
  distinct()
cat("Paulo 2015 — quiescent-enriched (≥2-fold):", nrow(paulo_q_up), "\n")

# ── Log-enriched (≥2-fold up in log phase = down during quiescence) ────────
si3_log <- read_excel(
  file.path(pub_dir, "Proteome/Paulo et al., 2015/NIHMS464126-supplement-3_si_003.xlsx"),
  sheet = "2X or more Up in LOG")
paulo_log_up <- si3_log %>%
  transmute(orf = Accession) %>%
  filter(is_orf(orf)) %>%
  distinct()
cat("Paulo 2015 — log-enriched (≥2-fold, i.e. decreasing in quiescence):",
    nrow(paulo_log_up), "\n")
```

# den Ridder et al. 2023

```{r denridder}
# Coverage comparison only (aerobic/anaerobic conditions, not quiescence)
dr_raw <- read_excel(
  file.path(pub_dir, "Proteome/den Ridder et al., 2023/mmc7.xlsx"),
  sheet = "All proteins")
denridder_detected <- dr_raw %>%
  transmute(accession = Accession,
            orf       = uniprot_to_orf(accession)) %>%
  filter(!is.na(orf)) %>%
  distinct(orf)
cat("den Ridder 2023 — detected:", nrow(denridder_detected), "\n")
```

# Coverage overlap (all detected proteins)

```{r coverage}
orf_gene <- ours_detected %>% filter(!is.na(gene)) %>%
  group_by(orf) %>% dplyr::slice(1) %>% ungroup()

coverage <- bind_rows(
  ours_detected      %>% transmute(orf, dataset = "ThisStudy"),
  paulo_detected     %>% transmute(orf, dataset = "Paulo2015"),
  denridder_detected %>% transmute(orf, dataset = "denRidder2023")) %>%
  distinct(orf, dataset) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  { if (!"ThisStudy"     %in% names(.)) mutate(., ThisStudy     = 0L) else . } %>%
  { if (!"Paulo2015"     %in% names(.)) mutate(., Paulo2015     = 0L) else . } %>%
  { if (!"denRidder2023" %in% names(.)) mutate(., denRidder2023 = 0L) else . } %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  arrange(desc(ThisStudy), desc(Paulo2015), desc(denRidder2023), orf)

write_csv(coverage, file.path(out_dir, "proteome_coverage.csv"))
cat("Coverage table:", nrow(coverage), "unique ORFs\n")

set_cov <- c("ThisStudy", "Paulo2015", "denRidder2023")
pdf(file.path(out_dir, "proteome_coverage_upset.pdf"), width = 7, height = 5)
print(upset(as.data.frame(coverage), sets = set_cov, keep.order = TRUE,
            order.by = "freq", mainbar.y.label = "Proteins (ORF)",
            sets.x.label = "Proteins per dataset"))
dev.off()
upset(as.data.frame(coverage), sets = set_cov, keep.order = TRUE,
      order.by = "freq", mainbar.y.label = "Proteins (ORF)")
```

# Starvation-regulation overlap (this study vs Paulo 2015)

```{r regulation}
# Four sets (all ORF-harmonized):
#   ThisStudy_up   = proteins ≥2-fold upregulated in WT at T30 vs T00 (C or N)
#   ThisStudy_down = proteins ≥2-fold downregulated in WT at T30 vs T00 (C or N)
#   Paulo_q_up     = proteins ≥2-fold enriched in quiescent cells (Paulo 2015)
#   Paulo_log_up   = proteins ≥2-fold enriched in log-phase cells (Paulo 2015)
#                    = proteins that decrease during quiescence in Paulo's study

regulation <- bind_rows(
  ours_up        %>% transmute(orf, dataset = "ThisStudy_up"),
  ours_down      %>% transmute(orf, dataset = "ThisStudy_down"),
  paulo_q_up     %>% transmute(orf, dataset = "Paulo_q_up"),
  paulo_log_up   %>% transmute(orf, dataset = "Paulo_log_up")) %>%
  distinct(orf, dataset) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from = dataset, values_from = present, values_fill = 0L) %>%
  { if (!"ThisStudy_up"   %in% names(.)) mutate(., ThisStudy_up   = 0L) else . } %>%
  { if (!"ThisStudy_down" %in% names(.)) mutate(., ThisStudy_down = 0L) else . } %>%
  { if (!"Paulo_q_up"     %in% names(.)) mutate(., Paulo_q_up     = 0L) else . } %>%
  { if (!"Paulo_log_up"   %in% names(.)) mutate(., Paulo_log_up   = 0L) else . } %>%
  left_join(orf_gene, by = "orf") %>%
  relocate(orf, gene) %>%
  arrange(desc(ThisStudy_up), desc(ThisStudy_down),
          desc(Paulo_q_up), desc(Paulo_log_up), orf)

write_csv(regulation, file.path(out_dir, "proteome_regulation_overlap.csv"))
cat("Regulation table:", nrow(regulation), "unique ORFs\n")

set_reg <- c("ThisStudy_up", "ThisStudy_down", "Paulo_q_up", "Paulo_log_up")
pdf(file.path(out_dir, "proteome_regulation_upset.pdf"), width = 8, height = 5)
print(upset(as.data.frame(regulation), sets = set_reg, keep.order = TRUE,
            order.by = "freq",
            mainbar.y.label = "Proteins (ORF)",
            sets.x.label   = "Proteins per set"))
dev.off()
upset(as.data.frame(regulation), sets = set_reg, keep.order = TRUE,
      order.by = "freq", mainbar.y.label = "Proteins (ORF)")
```

# Headline numbers

```{r headline}
cat("=== Coverage ===\n")
cat("This study detected:", sum(coverage$ThisStudy), "\n")
cat("Paulo 2015 detected:", sum(coverage$Paulo2015), "\n")
cat("den Ridder 2023 detected:", sum(coverage$denRidder2023), "\n")
cat("This study ∩ Paulo2015:", sum(coverage$ThisStudy & coverage$Paulo2015), "\n")
cat("This study ∩ denRidder2023:", sum(coverage$ThisStudy & coverage$denRidder2023), "\n")
cat("All three:", sum(coverage$ThisStudy & coverage$Paulo2015 & coverage$denRidder2023), "\n")

cat("\n=== Starvation regulation overlap (vs Paulo 2015) ===\n")
cat("This study upregulated (C or N, ≥2-fold at T30):", sum(regulation$ThisStudy_up), "\n")
cat("This study downregulated (C or N, ≥2-fold at T30):", sum(regulation$ThisStudy_down), "\n")
cat("Paulo quiescent-enriched:", sum(regulation$Paulo_q_up), "\n")
cat("Paulo log-enriched (down in quiescence):", sum(regulation$Paulo_log_up), "\n")

cat("\nUpregulated overlap (ThisStudy_up ∩ Paulo_q_up):",
    sum(regulation$ThisStudy_up & regulation$Paulo_q_up), "\n")
cat("Downregulated overlap (ThisStudy_down ∩ Paulo_log_up):",
    sum(regulation$ThisStudy_down & regulation$Paulo_log_up), "\n")
cat("Agreement fraction — up:",
    round(sum(regulation$ThisStudy_up & regulation$Paulo_q_up) /
            sum(regulation$ThisStudy_up), 3), "\n")
cat("Agreement fraction — down:",
    round(sum(regulation$ThisStudy_down & regulation$Paulo_log_up) /
            sum(regulation$ThisStudy_down), 3), "\n")

cat("\nTop upregulated proteins shared with Paulo quiescent-enriched set:\n")
up_shared <- regulation %>%
  filter(ThisStudy_up == 1, Paulo_q_up == 1) %>%
  left_join(fc_C %>% mutate(orf = uniprot_to_orf(accession)) %>%
              filter(!is.na(orf)) %>%
              transmute(orf, log2FC_C = log2FC),
            by = "orf") %>%
  left_join(fc_N %>% mutate(orf = uniprot_to_orf(accession)) %>%
              filter(!is.na(orf)) %>%
              transmute(orf, log2FC_N = log2FC),
            by = "orf") %>%
  arrange(desc(coalesce(log2FC_C, log2FC_N)))
print(up_shared %>% dplyr::select(orf, gene, log2FC_C, log2FC_N), n = 20)
```

```{r session}
sessionInfo()
```
