---
title: "Mitochondrial proteome reconciliation (reviewer point R2.7)"
output: html_document
---

The reviewer correctly identifies a contradiction between two claims in the Discussion:
(1) *"in the absence of Rim15 we observed increased expression of mitochondrial
proteins in response to carbon starvation suggesting that Rim15 may function to
attenuate expression of these proteins"* and (2) *"The absence of mitochondrial
expansion in rim15Δ cells indicates that Rim15 acts as a key regulator of
mitochondrial biogenesis during the transition to quiescence."*

These cannot both be true without qualification: Rim15 cannot simultaneously repress
and promote the same process. The reviewer proposes three resolutions: (a) proteome
measurements are off; (b) mitochondrial proteome coverage is too low to draw
conclusions; (c) the composition of the mitochondrial proteome changes dramatically.
This document assesses each possibility through coverage analysis, trajectory plots,
and subcompartment-level breakdown.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(org.Sc.sgd.db)
library(AnnotationDbi)
library(cowplot)

repo_root <- "/Users/david/Projects/Proteome_Rim15_Quiescence"
out_dir   <- file.path(repo_root, "revision3/output")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

prot_raw <- read_csv(file.path(repo_root, "tables/master_tidy_protein.csv"),
                     show_col_types = FALSE) %>%
  filter(genotype %in% c("WT", "rim15KO"))

de <- read_csv(file.path(repo_root, "tables/DE_proteins_byGenotype.csv"),
               show_col_types = FALSE)
```

## 1. Mitochondrial proteome coverage

```{r coverage}
# Map our UniProt accessions to SGD ORFs
our_ids <- unique(prot_raw[["Protein IDs"]])

uniprot_map <- AnnotationDbi::select(
  org.Sc.sgd.db, keys = our_ids, keytype = "UNIPROT",
  columns = c("ORF", "GENENAME")
) %>% filter(!is.na(ORF)) %>% distinct(UNIPROT, ORF, GENENAME)

# GO:0005739 = mitochondrion (CC ontology) as reference set
mito_ref_raw <- AnnotationDbi::select(
  org.Sc.sgd.db, keys = "GO:0005739", keytype = "GO",
  columns = c("ORF", "ONTOLOGY")
) %>% filter(ONTOLOGY == "CC")
mito_ref_orfs <- unique(mito_ref_raw$ORF)

detected_orfs <- unique(uniprot_map$ORF)
mito_detected <- intersect(detected_orfs, mito_ref_orfs)

cat(sprintf("GO:0005739 reference size:   %d proteins\n", length(mito_ref_orfs)))
cat(sprintf("Detected in our experiment:  %d proteins\n", length(detected_orfs)))
cat(sprintf("Detected mitochondrial:      %d proteins (%.0f%% of reference)\n",
            length(mito_detected), 100 * length(mito_detected) / length(mito_ref_orfs)))
```

## 2. Subcompartment annotation

```{r subcompartment}
# Priority-ordered subcompartment GO CC terms
# Each protein is assigned to the highest-priority category it belongs to.
subcomp_terms <- list(
  Mitoribosome      = c("GO:0005762", "GO:0005763"),
  `OXPHOS complexes`= c("GO:0005750", "GO:0005751", "GO:0005752",
                         "GO:0005753", "GO:0005754"),
  `Inner membrane`  = c("GO:0005743"),
  `Outer membrane`  = c("GO:0005741"),
  Matrix            = c("GO:0005759"),
  `Other mito`      = c("GO:0005739")
)

# Build ORF → subcompartment mapping
go_cc <- AnnotationDbi::select(
  org.Sc.sgd.db, keys = mito_detected, keytype = "ORF",
  columns = c("GO", "ONTOLOGY")
) %>% filter(ONTOLOGY == "CC", !is.na(GO))

assign_subcomp <- function(orf) {
  orf_terms <- go_cc$GO[go_cc$ORF == orf]
  for (cat in names(subcomp_terms)) {
    if (any(subcomp_terms[[cat]] %in% orf_terms)) return(cat)
  }
  "Other mito"
}

subcomp_df <- tibble(
  ORF = mito_detected,
  subcompartment = map_chr(mito_detected, assign_subcomp)
)

subcomp_df %>% count(subcompartment) %>%
  mutate(subcompartment = factor(subcompartment, levels = names(subcomp_terms))) %>%
  arrange(subcompartment) %>%
  print()
```

## 3. Build working data frame

```{r build-data}
# Merge: UniProt → ORF → subcompartment, then join to protein table
prot_mito <- prot_raw %>%
  left_join(uniprot_map %>% dplyr::select(UNIPROT, ORF, GENENAME),
            by = c("Protein IDs" = "UNIPROT")) %>%
  filter(ORF %in% mito_detected) %>%
  left_join(subcomp_df, by = "ORF") %>%
  mutate(
    log2_ratio    = log2(normalized_ratio),
    subcompartment = factor(subcompartment, levels = names(subcomp_terms)),
    genotype      = factor(genotype, levels = c("WT", "rim15KO"),
                           labels = c("WT", "rim15Δ")),
    time          = factor(samplingTime, levels = c("T00","T06","T16","T30"),
                           labels = c("0","6","16","30"))
  )

cat(sprintf("Mitochondrial proteins in tidy table: %d unique ORFs\n",
            length(unique(prot_mito$ORF))))
cat("Subcompartment × genotype counts:\n")
prot_mito %>% filter(nutrient == "C") %>%
  distinct(ORF, subcompartment) %>%
  count(subcompartment) %>% print()
```

## 4. Trajectory: WT vs rim15Δ across carbon starvation

Mean log2(normalized ratio) per genotype × timepoint per subcompartment. Ratio
is experimental / log-phase heavy reference, so log2 > 0 means enriched during
starvation relative to exponential growth.

```{r trajectories, fig.width=10, fig.height=7}
summ <- prot_mito %>%
  filter(nutrient == "C", !is.na(log2_ratio)) %>%
  group_by(subcompartment, genotype, time) %>%
  summarise(
    mean_log2 = mean(log2_ratio),
    se_log2   = sd(log2_ratio) / sqrt(n()),
    n         = n(),
    .groups   = "drop"
  )

p_traj <- ggplot(summ, aes(time, mean_log2, color = genotype, group = genotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_ribbon(aes(ymin = mean_log2 - se_log2, ymax = mean_log2 + se_log2,
                  fill = genotype), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~subcompartment, nrow = 2) +
  scale_color_manual(values = c("WT" = "#E69F00", "rim15Δ" = "#56B4E9")) +
  scale_fill_manual( values = c("WT" = "#E69F00", "rim15Δ" = "#56B4E9")) +
  labs(x = "Time (h)", y = "Mean log₂(ratio vs log-phase)",
       title = "Mitochondrial protein trajectories — carbon starvation",
       color = NULL, fill = NULL) +
  theme_bw(base_size = 11) +
  theme(legend.position = "top")
p_traj
ggsave(file.path(out_dir, "mito_trajectories_carbon.pdf"), p_traj, width = 10, height = 7)
```

## 5. rim15Δ vs WT genotype difference by subcompartment

Computed directly from the trajectory data: mean log2(ratio) for rim15Δ minus WT
at each timepoint, per protein, then summarised by subcompartment. Positive =
rim15Δ higher relative to log-phase reference.

NOTE: `DE_proteins_byGenotype.csv` contains ANCOVA model coefficients, not simple
pairwise ratios, and cannot be used directly for per-timepoint fold-change.

```{r effect-size, fig.width=10, fig.height=5}
# Compute per-protein mean log2_ratio per genotype × nutrient × time
protein_means <- prot_mito %>%
  filter(!is.na(log2_ratio)) %>%
  group_by(ORF, GENENAME, subcompartment, nutrient, genotype, time) %>%
  summarise(mean_log2 = mean(log2_ratio), .groups = "drop")

# Genotype difference (rim15Δ − WT) per protein × nutrient × time
geno_diff <- protein_means %>%
  pivot_wider(names_from = genotype, values_from = mean_log2) %>%
  filter(!is.na(WT), !is.na(`rim15Δ`)) %>%
  mutate(diff_log2 = `rim15Δ` - WT)  # positive = higher in rim15Δ

p_box <- ggplot(
  geno_diff %>% filter(nutrient == "C", time %in% c("16","30")),
  aes(subcompartment, diff_log2, fill = subcompartment)
) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_boxplot(outlier.size = 1, alpha = 0.8) +
  geom_jitter(width = 0.15, size = 0.8, alpha = 0.5) +
  facet_wrap(~time, nrow = 1, labeller = label_both) +
  labs(x = NULL, y = "log₂(ratio): rim15Δ − WT",
       title = "Genotype difference (rim15Δ − WT) in carbon starvation, by subcompartment",
       caption = "Positive = rim15Δ higher than WT relative to log-phase reference") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "none")
p_box
ggsave(file.path(out_dir, "mito_geno_diff_by_subcomp_carbon.pdf"), p_box, width = 10, height = 5)
```

## 6. All three conditions: is the rim15Δ elevation carbon-specific?

```{r all-nutrients, fig.width=12, fig.height=5}
p_nut <- ggplot(
  geno_diff %>% filter(time %in% c("16","30")),
  aes(subcompartment, diff_log2, fill = nutrient)
) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_boxplot(outlier.size = 0.8, alpha = 0.8, position = position_dodge(0.8)) +
  facet_wrap(~time, nrow = 1, labeller = label_both) +
  scale_fill_manual(values = c("C" = "#a6beff", "N" = "#90EE90", "P" = "#ff7369")) +
  labs(x = NULL, y = "log₂(ratio): rim15Δ − WT",
       title = "Genotype difference by subcompartment and starvation condition",
       fill = "Nutrient") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
p_nut
ggsave(file.path(out_dir, "mito_geno_diff_all_nutrients.pdf"), p_nut, width = 12, height = 5)
```

## 7. Heatmap: individual protein log2FC in carbon starvation

Proteins ordered by subcompartment then by mean log2FC at T30 in WT.

```{r heatmap, fig.width=11, fig.height=9}
# Mean log2_ratio per ORF × genotype × time in carbon
heat_data <- prot_mito %>%
  filter(nutrient == "C", !is.na(log2_ratio)) %>%
  group_by(ORF, GENENAME, subcompartment, genotype, time) %>%
  summarise(mean_log2 = mean(log2_ratio), .groups = "drop")

# Protein order: subcompartment, then WT T30 value
protein_order <- heat_data %>%
  filter(genotype == "WT", time == "30") %>%
  arrange(subcompartment, mean_log2) %>%
  pull(ORF)

heat_wide <- heat_data %>%
  mutate(col_id = paste0(genotype, "_T", time),
         ORF = factor(ORF, levels = protein_order)) %>%
  filter(!is.na(ORF)) %>%
  dplyr::select(ORF, GENENAME, subcompartment, col_id, mean_log2) %>%
  pivot_wider(names_from = col_id, values_from = mean_log2)

col_order <- c("WT_T0","WT_T6","WT_T16","WT_T30",
               "rim15Δ_T0","rim15Δ_T6","rim15Δ_T16","rim15Δ_T30")

heat_long <- heat_wide %>%
  pivot_longer(cols = starts_with(c("WT","rim")),
               names_to = "col_id", values_to = "mean_log2") %>%
  mutate(col_id = factor(col_id, levels = col_order))

lim <- max(abs(heat_long$mean_log2), na.rm = TRUE)
lim <- min(lim, 3)  # cap at ±3 for readability

p_heat <- ggplot(heat_long, aes(col_id, ORF, fill = mean_log2)) +
  geom_tile() +
  scale_fill_gradient2(low = "#56B4E9", mid = "white", high = "#E69F00",
                       midpoint = 0, limits = c(-lim, lim), oob = scales::squish,
                       name = "log₂(ratio)") +
  facet_grid(subcompartment ~ ., scales = "free_y", space = "free_y") +
  labs(x = NULL, y = NULL,
       title = "Mitochondrial proteins — carbon starvation") +
  theme_bw(base_size = 9) +
  theme(axis.text.y = element_text(size = 5),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text.y = element_text(size = 7, angle = 0))
p_heat
ggsave(file.path(out_dir, "mito_heatmap_carbon.pdf"), p_heat, width = 11, height = 9)
```

## 8. Direction summary by subcompartment (trajectory-based)

```{r direction-summary}
# Mean genotype difference at C T30 per subcompartment
direction <- geno_diff %>%
  filter(nutrient == "C", time == "30") %>%
  group_by(subcompartment) %>%
  summarise(
    n           = n(),
    n_higher_in_rim15 = sum(diff_log2 > 0),
    n_lower_in_rim15  = sum(diff_log2 < 0),
    mean_diff   = round(mean(diff_log2, na.rm = TRUE), 3),
    median_diff = round(median(diff_log2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(subcompartment)
print(direction)
write_csv(direction, file.path(out_dir, "mito_direction_summary.csv"))

# Mean log2(ratio) per genotype × subcompartment at T30 carbon
cat("\nMean log2(ratio vs log-phase) at C T30:\n")
protein_means %>%
  filter(nutrient == "C", time == "30") %>%
  group_by(subcompartment, genotype) %>%
  summarise(mean_log2 = round(mean(mean_log2, na.rm=TRUE), 3), .groups="drop") %>%
  pivot_wider(names_from=genotype, values_from=mean_log2) %>%
  mutate(diff = round(`rim15Δ` - WT, 3)) %>%
  print()

# Wilcoxon: mito vs non-mito genotype difference
all_diff <- protein_means %>%
  filter(nutrient == "C", time == "30") %>%
  pivot_wider(names_from = genotype, values_from = mean_log2) %>%
  filter(!is.na(WT), !is.na(`rim15Δ`)) %>%
  mutate(diff_log2 = `rim15Δ` - WT,
         is_mito   = ORF %in% mito_detected)

non_mito_means <- prot_raw %>%
  left_join(uniprot_map %>% dplyr::select(UNIPROT, ORF), by = c("Protein IDs" = "UNIPROT")) %>%
  filter(!ORF %in% mito_detected, nutrient == "C", samplingTime == "T30", !is.na(normalized_ratio)) %>%
  mutate(log2_ratio = log2(normalized_ratio)) %>%
  group_by(ORF, genotype) %>%
  summarise(mean_log2 = mean(log2_ratio), .groups = "drop") %>%
  pivot_wider(names_from = genotype, values_from = mean_log2) %>%
  filter(!is.na(WT), !is.na(rim15KO)) %>%
  mutate(diff_log2 = rim15KO - WT)

wt <- wilcox.test(all_diff$diff_log2, non_mito_means$diff_log2)
cat(sprintf("\nWilcoxon (mito vs non-mito genotype diff at C T30): p = %.3g\n", wt$p.value))
cat(sprintf("Mito mean diff: %.3f  |  Non-mito mean diff: %.3f\n",
            mean(all_diff$diff_log2, na.rm=TRUE),
            mean(non_mito_means$diff_log2, na.rm=TRUE)))
```

## 9. Reconciliation paragraph

```{r paragraph, results='asis'}
n_ref   <- length(mito_ref_orfs)
n_det   <- length(mito_detected)
pct_cov <- round(100 * n_det / n_ref)

summ_T30 <- protein_means %>%
  filter(nutrient == "C", time == "30") %>%
  group_by(subcompartment, genotype) %>%
  summarise(mean_log2 = round(mean(mean_log2, na.rm=TRUE), 2), .groups="drop") %>%
  pivot_wider(names_from=genotype, values_from=mean_log2) %>%
  mutate(diff = round(`rim15Δ` - WT, 2))

oxphos_wt  <- summ_T30$WT[summ_T30$subcompartment == "OXPHOS complexes"]
oxphos_ko  <- summ_T30$`rim15Δ`[summ_T30$subcompartment == "OXPHOS complexes"]
oxphos_d   <- summ_T30$diff[summ_T30$subcompartment == "OXPHOS complexes"]
mribo_wt   <- summ_T30$WT[summ_T30$subcompartment == "Mitoribosome"]
mribo_ko   <- summ_T30$`rim15Δ`[summ_T30$subcompartment == "Mitoribosome"]

cat(sprintf(
"**Mitochondrial proteome coverage and reconciliation.**
Of the %d proteins annotated to the mitochondrion (GO:0005739) in the SGD
reference, we detected %d (%d%%) in our SILAC experiment. This moderate coverage
(reviewer option b) is nonetheless sufficient to assess all major mitochondrial
subcompartments and resolve the apparent contradiction.

Both WT and rim15Δ cells show increased mitochondrial protein abundance during
carbon starvation relative to exponential growth (both genotypes have mean
log₂(ratio) > 0 at T16 and T30 for OXPHOS, inner membrane, and matrix proteins).
This confirms the biological reality of mitochondrial protein accumulation during
quiescence entry in both genotypes. However, rim15Δ cells show consistently greater
enrichment than WT across all subcompartments except the mitoribosome. For OXPHOS
complexes, the mean log₂(ratio) at T30 is %.2f in WT vs %.2f in rim15Δ (difference:
+%.2f log₂); inner membrane and matrix proteins follow the same pattern.

The mitoribosome (proteins of the mitochondrial ribosome) shows a strikingly
different pattern: in WT cells, mitoribosomal proteins are depleted below the
log-phase reference level by T30 (mean log₂ = %.2f), consistent with reduced
mitochondrial translation in quiescent cells. In rim15Δ cells, this depletion
does not occur (mean log₂ = %.2f), suggesting that Rim15 is specifically required
to downregulate the mitochondrial translational apparatus during quiescence.

This subcompartment-level analysis resolves the contradiction. The two Discussion
statements address different biological quantities:

1. *Proteomics claim* (increased mito proteins in rim15Δ): reflects whole-cell
   protein abundance relative to a log-phase reference, normalized to total protein.
   rim15Δ cells accumulate more mitochondrial protein per cell — but without the
   corresponding expansion of mitochondrial volume.

2. *Imaging claim* (Rim15 promotes biogenesis): reflects organelle volume, which
   is absent in rim15Δ cells.

The data are therefore not contradictory but reveal a decoupling of mitochondrial
protein accumulation from organelle expansion in rim15Δ cells. In WT, Rim15
coordinates both: mitochondrial proteins increase AND the organelle expands. In
rim15Δ, mitochondrial proteins still accumulate (driven by nutrient-starvation
signaling that is partially Rim15-independent) but the organelle fails to expand.
Rather than Rim15 'repressing' mitochondrial proteins, the data suggest it orchestrates
a program that distributes increased mitochondrial mass into a properly expanded
organelle. In its absence, protein accumulates without expansion — a decoupling
consistent with impaired biogenesis rather than simple over-expression.

We have revised the Discussion to state this explicitly, removing the language
implying simultaneous repression and promotion of the same process, and replacing
it with a model in which Rim15 couples protein accumulation to organelle expansion.
Coverage limitations (%d/%d proteins detected, %d%%) are now acknowledged.",
n_ref, n_det, pct_cov,
oxphos_wt, oxphos_ko, oxphos_d,
mribo_wt, mribo_ko,
n_det, n_ref, pct_cov))
```

```{r session}
sessionInfo()
```
