---
title: "Combined-genotype PCA of the proteome (reviewer point R2.6)"
output: html_document
---

Reviewer 2 noted that in the per-genotype PCA (Fig S2C) PC2 appears to separate
replicates on carbon, with one WT carbon replicate offset, and asked whether
genotype separates more than replicate scatter when WT and rim15KO are analyzed
together. Here we run PCA on the pooled WT + rim15KO normalized SILAC ratios,
color by genotype / nutrient / time / replicate, and check sensitivity of the
conclusions to the suspect WT carbon replicate.

```{r setup, message=FALSE}
library(tidyverse)

ratios <- read_csv("./tables/normed_ratio_perseus.csv")

dir.create("./revision3/output", showWarnings = FALSE, recursive = TRUE)
```

```{r prepare-matrix}
# Sample columns are named genotype_nutrient_time_replicate, e.g. WT_C_T06_R2.
sample_cols <- grep("^(WT|rim15KO)_", names(ratios), value = TRUE)

# Matrix: samples in rows (like observations), proteins in columns.
# Keep proteins quantified in every sample so PCA needs no imputation
# (mirrors the complete-case approach used in the per-genotype PCA).
mat <- ratios %>%
  select(`Protein IDs`, all_of(sample_cols)) %>%
  column_to_rownames("Protein IDs") %>%
  as.matrix()

complete <- mat[rowSums(is.na(mat)) == 0, , drop = FALSE]
message(nrow(complete), " proteins complete across all ", ncol(complete), " samples")

# log2 the ratios before PCA
pca_in <- t(log2(complete))

meta <- tibble(sample = rownames(pca_in)) %>%
  separate(sample, into = c("genotype", "nutrient", "time", "replicate"),
           sep = "_", remove = FALSE)
```

```{r pca}
pc <- prcomp(pca_in, center = TRUE, scale. = FALSE)
var_expl <- 100 * pc$sdev^2 / sum(pc$sdev^2)

scores <- as_tibble(pc$x[, 1:4], rownames = "sample") %>% left_join(meta, by = "sample")

plot_pca <- function(color_by) {
  ggplot(scores, aes(PC1, PC2, color = .data[[color_by]], shape = genotype)) +
    geom_point(size = 3) +
    labs(x = sprintf("PC1 (%.1f%%)", var_expl[1]),
         y = sprintf("PC2 (%.1f%%)", var_expl[2]),
         color = color_by) +
    theme_bw()
}

p <- cowplot::plot_grid(plot_pca("genotype"), plot_pca("nutrient"),
                        plot_pca("time"), plot_pca("replicate"), nrow = 2)
p
ggsave("./revision3/output/combined_pca.pdf", p, width = 10, height = 8)

# Which PC captures genotype? Correlate PCs with genotype indicator.
scores %>%
  mutate(gt = as.numeric(genotype == "WT")) %>%
  summarise(across(starts_with("PC"), ~ cor(.x, gt))) %>%
  pivot_longer(everything(), names_to = "PC", values_to = "cor_with_genotype") %>%
  arrange(desc(abs(cor_with_genotype)))
```

```{r outlier-sensitivity}
# Identify the offset WT carbon replicate on PC2 (per Fig S2C), then re-run the
# PCA without that replicate and confirm the genotype/nutrient/time structure
# is unchanged.
scores %>% filter(genotype == "WT", nutrient == "C") %>%
  select(sample, PC1, PC2) %>% arrange(PC2)

# TODO once the outlier sample is identified from the table above:
# drop it, re-run prcomp, and compare loadings/score structure.
```

```{r session}
sessionInfo()
```
