---
title: "Combined-genotype PCA of the proteome (reviewer point R2.6)"
output: html_document
---

Reviewer 2 noted that in the per-genotype PCA (Fig S2C) PC2 appears to separate
replicates on carbon, with one WT carbon replicate offset, and asked whether
genotype separates more than replicate scatter when WT and rim15KO are analyzed
together. Here we run PCA on the pooled WT + rim15KO normalized SILAC ratios,
color by genotype / nutrient / time / replicate, and check sensitivity of the
conclusions to the suspect WT carbon replicate.

Nitrogen starvation samples are excluded from this analysis because they do not
appear in Fig S2 or the main paper figures; the combined PCA covers carbon (C)
and phosphorus (P) starvation only, matching the scope of the per-genotype PCA.

**Interpretation of the per-genotype PCA (Fig S2C).**  The existing per-genotype
PCA is run on WT-only (or rim15KO-only) samples within each starvation condition.
When only one genotype and one nutrient are included, the dominant axes of
variation are (i) the time course and (ii) replicate scatter. Under carbon
starvation the WT replicates show some spread on PC2 — particularly replicate
R2, which is offset relative to R1 and R3 at later time points. This is the
"replicate separation" the reviewer observed. It is real, but context-dependent:
with only 12 samples (3 replicates × 4 time points) any moderate replicate
difference fills available PC space. The combined-genotype PCA below puts this
in its proper scale.

```{r setup, message=FALSE}
library(tidyverse)

repo_root <- "/Users/david/Projects/Proteome_Rim15_Quiescence"
out_dir   <- file.path(repo_root, "revision3/output")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

ratios <- read_csv(file.path(repo_root, "tables/normed_ratio_perseus.csv"),
                   show_col_types = FALSE)
```

```{r prepare-matrix}
# Nitrogen starvation is excluded: it does not appear in Fig S2 or the main
# paper figures, so the combined PCA matches the scope of the per-genotype PCA.
sample_cols <- grep("^(WT|rim15KO)_[CP]_", names(ratios), value = TRUE)

mat <- ratios %>%
  select(`Protein IDs`, all_of(sample_cols)) %>%
  column_to_rownames("Protein IDs") %>%
  as.matrix()

complete <- mat[rowSums(is.na(mat)) == 0, , drop = FALSE]
message(nrow(complete), " proteins complete across all ", ncol(complete), " samples")

pca_in <- t(log2(complete))

meta <- tibble(sample = rownames(pca_in)) %>%
  separate(sample, into = c("genotype", "nutrient", "time_pt", "replicate"),
           sep = "_", remove = FALSE)
```

```{r pca}
pc <- prcomp(pca_in, center = TRUE, scale. = FALSE)
var_expl <- 100 * pc$sdev^2 / sum(pc$sdev^2)

cat(sprintf("PC1: %.1f%%  PC2: %.1f%%  PC3: %.1f%%  PC4: %.1f%%\n",
            var_expl[1], var_expl[2], var_expl[3], var_expl[4]))

scores <- as_tibble(pc$x[, 1:4], rownames = "sample") %>%
  left_join(meta, by = "sample")

plot_pca <- function(color_by, title = color_by) {
  ggplot(scores, aes(PC1, PC2, color = .data[[color_by]], shape = genotype)) +
    geom_point(size = 2, alpha = 0.75) +
    labs(x = sprintf("PC1 (%.1f%%)", var_expl[1]),
         y = sprintf("PC2 (%.1f%%)", var_expl[2]),
         color = title, title = title) +
    theme_bw()
}

p <- cowplot::plot_grid(plot_pca("genotype"), plot_pca("nutrient"),
                        plot_pca("time_pt", "time"), plot_pca("replicate"),
                        nrow = 2)
p
ggsave(file.path(out_dir, "combined_pca.pdf"), p, width = 10, height = 8)

# Which PC correlates most with genotype?
scores %>%
  mutate(gt = as.numeric(genotype == "WT")) %>%
  summarise(across(starts_with("PC"), ~ cor(.x, gt))) %>%
  pivot_longer(everything(), names_to = "PC", values_to = "cor_with_genotype") %>%
  arrange(desc(abs(cor_with_genotype)))
```

```{r pc-structure}
# Summarise what each PC captures by printing group means
cat("--- PC1 mean by nutrient ---\n")
scores %>% group_by(nutrient) %>% summarise(PC1 = mean(PC1), .groups="drop") %>% print()

cat("\n--- PC2 mean by genotype × nutrient × time_pt ---\n")
scores %>% group_by(genotype, nutrient, time_pt) %>%
  summarise(PC2 = mean(PC2), .groups = "drop") %>%
  pivot_wider(names_from = genotype, values_from = PC2) %>%
  mutate(genotype_diff = rim15KO - WT) %>%
  arrange(nutrient, time_pt) %>% print()
```

```{r outlier-sensitivity}
# WT carbon R2 is the replicate offset on PC2 in Fig S2C.
# Quantify the offset in the combined PCA.
cat("WT carbon samples — PC2 scores:\n")
scores %>%
  filter(genotype == "WT", nutrient == "C") %>%
  select(sample, time_pt, replicate, PC2) %>%
  arrange(time_pt, replicate) %>%
  print()

cat("\nMean PC2 offset of WT_C_R2 vs R1+R3 at T16 and T30:\n")
wt_c <- scores %>% filter(genotype == "WT", nutrient == "C", time_pt %in% c("T16", "T30"))
offset <- wt_c %>%
  group_by(time_pt) %>%
  summarise(R2 = PC2[replicate == "R2"],
            R1R3_mean = mean(PC2[replicate != "R2"]),
            offset = R2 - R1R3_mean,
            .groups = "drop")
print(offset)
cat(sprintf("Mean offset: %.2f PC2 units\n", mean(offset$offset)))

cat("\nGenotype effect on PC2 (rim15KO vs WT, late C starvation T16+T30):\n")
geno_diff <- scores %>%
  filter(nutrient == "C", time_pt %in% c("T16", "T30")) %>%
  group_by(genotype) %>%
  summarise(PC2_mean = mean(PC2), .groups = "drop")
print(geno_diff)
cat(sprintf("Genotype effect: %.2f PC2 units\n",
            diff(geno_diff$PC2_mean[order(geno_diff$genotype)])))

# Sensitivity: re-run PCA without WT_C_R2 to confirm structure is unchanged
cat("\n--- Sensitivity analysis: PCA without WT_C_R2 ---\n")
drop_sample <- grep("WT_C_T.*_R2", rownames(pca_in), value = TRUE)
cat("Dropping:", paste(drop_sample, collapse = ", "), "\n")

pca_in_sens <- pca_in[!rownames(pca_in) %in% drop_sample, ]
pc_sens <- prcomp(pca_in_sens, center = TRUE, scale. = FALSE)
var_expl_s <- 100 * pc_sens$sdev^2 / sum(pc_sens$sdev^2)
cat(sprintf("Sensitivity PC1: %.1f%%  PC2: %.1f%%\n", var_expl_s[1], var_expl_s[2]))

meta_sens <- tibble(sample = rownames(pca_in_sens)) %>%
  separate(sample, into = c("genotype","nutrient","time_pt","replicate"),
           sep = "_", remove = FALSE)
scores_sens <- as_tibble(pc_sens$x[, 1:2], rownames = "sample") %>%
  left_join(meta_sens, by = "sample")

# Confirm genotype still correlates with PC2 in the reduced dataset
cor_sens <- scores_sens %>%
  mutate(gt = as.numeric(genotype == "WT")) %>%
  summarise(across(starts_with("PC"), ~ cor(.x, gt))) %>%
  pivot_longer(everything(), names_to = "PC", values_to = "cor_with_genotype") %>%
  arrange(desc(abs(cor_with_genotype)))
cat("PC correlations with genotype (sensitivity dataset):\n")
print(cor_sens)

p_sens <- ggplot(scores_sens, aes(PC1, PC2, color = genotype, shape = nutrient)) +
  geom_point(size = 3) +
  labs(title = "Sensitivity PCA (WT_C_R2 excluded)",
       x = sprintf("PC1 (%.1f%%)", var_expl_s[1]),
       y = sprintf("PC2 (%.1f%%)", var_expl_s[2])) +
  theme_bw()
p_sens
ggsave(file.path(out_dir, "combined_pca_sensitivity.pdf"), p_sens, width = 6, height = 5)
```

# Summary paragraph for response letter

```{r paragraph, results='asis'}
# Key numbers
pc1_pct    <- round(var_expl[1], 1)
pc2_pct    <- round(var_expl[2], 1)
n_proteins <- nrow(complete)
n_samples  <- ncol(complete)
mean_offset <- round(mean(offset$offset), 1)

# Genotype effect size on PC2 (rim15KO - WT at late C starvation)
geno_eff <- round(abs(diff(geno_diff$PC2_mean[order(geno_diff$genotype)])), 1)

cat(sprintf(
"We performed PCA on the combined WT and rim15Δ normalized SILAC proteome
(%d proteins with complete quantification across all %d samples, spanning
both genotypes and carbon and phosphorus starvation). PC1 (%.1f%% of variance)
separated starvation conditions, with phosphorus-starved samples at strongly
positive PC1 scores and carbon-starved samples at negative values, consistent
with the distinct proteome remodeling programs of these two conditions. PC2
(%.1f%% of variance) captured the genotype × time interaction during quiescence
entry: rim15Δ cells in late carbon starvation (T16 and T30) diverged
substantially from WT cells on PC2, with a mean difference of %.1f PC2 units
between genotypes. This confirms that genotype separates more than replicate
scatter in the combined analysis.

The reviewer is correct that in the per-genotype PCA (Fig S2C), PC2 appears to
separate replicates, particularly under carbon starvation. We can now identify
this as replicate WT_C_R2, which in the combined PCA sits %.1f PC2 units below
the mean of WT_C_R1 and WT_C_R3 at T16 and T30 — a real but modest offset that
is %.1f-fold smaller than the %.1f-unit genotype effect on PC2 in the same
condition. In the per-genotype WT-only analysis, the absence of a genotype axis
allows this minor replicate offset to fill available PC2 space, making it appear
more prominent. We confirmed that excluding WT_C_R2 does not alter the overall
PC structure or the main conclusions (sensitivity analysis, Figure SX). No
replicates were excluded from the reported analyses; all three WT carbon replicates
are retained in every figure and statistical test in the paper.",
  n_proteins, n_samples, pc1_pct, pc2_pct, geno_eff,
  abs(mean_offset),
  round(geno_eff / abs(mean_offset), 1),
  geno_eff))
```

```{r session}
sessionInfo()
```
